---
title: "GOA rex sole assessment instructions"
author: "Carey McGilliard"
format: html
editor: visual
---

# GOA rex sole assessment instructions

## Obtaining GOA rex sole assessment data inputs

The assessment uses the following data:

### Catch biomass

- One area, as there is only fishing in the western-central GOA and nearly no fishing in the Eastern GOA.

```         
- Download directly from AKFIN Answers ([https://akfinbi.psmfc.org/](https://akfinbi.psmfc.org/bi-security-login/login.jsp?msi=false&redirect=L2FuYWx5dGljcy9zYXcuZGxsP2JpZWVob21lJnN0YXJ0UGFnZT0xJmhhc2g9Y0MtckpJLUVnQUhzUlBtMDM5TWZBamZxdkctT0oydlhlc0VON0h2azlyeFFOSlZRZkNMTVlHYWRNdXRTcmF1Wg==))

  - check boxes for both rex sole and GOA rex sole and grab all years, compare to the previous assessment.
```

### Fishery comp data

- Fishery data are pulled using the goa_rex example in the sampler repo and include fishery age comp and fishery length comp

- Plan is to update fishery sample size methods when Pete's new repo for processing fishery data is done, which is post-2026 assessment

- To do for 2026: write query in the sampler repo to pull the sample sizes for the observer data if they are not output from sampler

- Before 2025:

  - Fishery ages: (this was a criticism of the CIE review - it uses raw lengths rather than extrapolated lengths to include the port data): source("C:\\GitProjects\\newsbss\\Inputs_PieceByPiece\\Get_Fishery_Ages\\Get_Fishery_Ages_Port_or_Haul_noStratifying_noExtrapLens.R")

  - Fishery lengths: Before 2025 these came from: source("C:\\GitProjects\\newsbss\\Get_Fishery_Lengths\\Get_Fishery_Lengths_With_Extrapolated_Number.R")

  - Sample size (Nsamp) for Fishery lengths (and tables for the SAFE): update directories and source: source("C:\\GitProjects\\newsbss\\Get_Fishery_Lengths\\Get_Fishery_Length_HaulStats_With_NORPAQ.R")

### Survey Data

- For 2025/2026: see C:/GitProjects/pull_gap_products/examples/goa_rex.qmd which has the code used for pulling all of the survey data and comparing to the 2021 assessment

- Before 2025 (much of this code will not work as it pre-dates gap_products):

  - Survey lengths: source("C:\\GitProjects\\newsbss\\Get_GOA_Survey_Length_and_Age_Comp\\Get_GOA_Length_Comps_by_Area.R")

  - Sample sizes for lengths and ages: source (for nsamp for survey lengths and for ages), by area (made with Wayne Palsson's help) source("C:\\GitProjects\\newsbss\\Get_GOA_Survey_Length_and_Age_Comp\\Get_Survey_Length_Age_Stats_By_Area.R")

  - Note that there are lines for reading in the data from .Rdata files, but below this are the queries for getting the data from SQL and then saving those same .Rdata files #This file also gets the number of hauls. #Don't forget to add in columns for ages that don't appear in the data - this isn't automated. #Run with #FmpArea = "600 and 699" #ByGear = 0 #ByPortOrHaul = 0 #LengthBins = seq(9,65,2) #AgeBins = seq(1,20,1)

  - Survey conditional age-at-lengths #do the SQL query and write the data necessary to read in to the next step here (which will also plot a bunch of age-length ggplots): source("C:\\GitProjects\\newsbss\\Inputs_PieceByPiece\\Get_GOA_Survey_Length_and_Age_Comp\\Get_Survey_Length_Age_and_Plot.R")

  - Run this to get the conditional age-at-length data roughly formatted for assessment input source("C:\\GitProjects\\newsbss\\Inputs_PieceByPiece\\Get_GOA_Survey_Length_and_Age_Comp\\GET_GOA_CONDITIONAL_AGE_AT_LENGTH.R") #run it for "eastern" and for "noneastern" separately #len_bins = seq(9,65,2) #max_age = 20 #eastern = "eastern" then do eastern = "noneastern" #default is 0

  - GOA marginal age comps were tricky because they are not produced by the survey group by area (or by Western-Central and Eastern), otherwise they would be included as ghost comps. (Pete and Ben were trying to resolve this after 2021 using area-specific age-length keys for afscISS and surveyISS. However, I couldn't get their caal data to match mine without any bootstrapping in 2025/2026)

## Running the ss3 assessment model and producing model plots

This uses the newsbss repo (which used to also be used for data inputs but is not anymore). r4ss is run here: C:/GitProjects/newsbss/Process_SS_Output/SS3toProjectTables.R. Update the fields at the top. This messy code does the following:

- runs r4ss

- makes r4ss plots

- makes additional plots

- calculates updated data weighting

- writes some parameter tables for the document (just simple .csv files)

- writes the data files for doing projections

## Projections

#Before running the projection part of the code below do 2 manual steps: #1. update AK-projection-simulation repo (pull updates) #2. update the projection input files aside from the main (ForProjections) data file (these are in: #C:\Users\carey.mcgilliard\Work\FlatfishAssessments\ProjectionModel\RexTemplate) #Updating requires updated catches and updating the number of years of projections (14 for a full assessment year with a new data) and new catches and check \# years of catches input in proj input files #Probably worthwhile to check into whether it's better to run Jim's R code in AK-projection-simulation repo?? but the code below is for SS output) #Carey can never get Jim's newest projection code to run.

#To: #1. plot SS results #2. Find Francis or McAllister and Ianelli Data weighting update values (for variance adjustments) #3. Plot fishery selex and maturity comparison for rex (as this was what made bio ref pts unreliable in the single area model - fishery selex way to right of maturity) #4. Plot growth data and estimated growth curves for both areas (uses AgeLength.csv, which was made earlier in newsbss/Inputs_PieceByPiece/Get_Survey_Length_Age_and_Plot.R) #5. Write .csv files for making parameter reporting tables #6. Write the projection input file to run folder (Careful: be sure to check over the projection file, especially female and male maturity) #(note writing directly to the run folder is meant to help prevent reporting an exec summary table for the wrong run and also helps when comparing projections for multiple runs) #7. Run projections: source("C:\\GitProjects\\newsbss\\Process_SS_Output\\SS3toProjectTables.R")

#To make the executive summary table and harvest spec tables and do the required phase plot: source(C:\GitProjects\newsbss\Process\_SS_Output\ExecSummaryTable\_Phase_Plot_HarvestRecsTable.R)

#To run retrospectives: #Make a new folder (Like Run_8_Retro), run the following (and this will plot the retros too): #change file paths and run folders source("C:\\GitProjects\\newsbss\\Process_SS_Output\\RunRetrospectives.R")

#Compare any alternative runs using: #change file paths and run folders source("C:\\GitProjects\\newsbss\\Process_SS_Output\\CompareSS3Runs.R")

## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
