---
title: "GOA rex sole assessment instructions"
author: "Carey McGilliard"
format: html
editor: visual
---

# GOA rex sole assessment instructions

## Read in libraries

```{r}

library(devtools)
devtools::install_github("afsc-assessments/afscISS")

#library(devtools)
#devtools::install_github("BenWilliams-NOAA/surveyISS")
install.packages("C:/GitProjects/surveyISS", repos = NULL, type = "source")

devtools::install_github("careymcg/AreaGrowthRex")
devtools::install_github("afsc-assessments/rema", dependencies = TRUE, build_vignettes = TRUE)
devtools::install_github(repo = "afsc-assessments/spmR", dependencies = TRUE, build_vignettes = TRUE)

if (!requireNamespace("pacman", quietly = TRUE)) {
  install.packages("pacman")
}

newlibs <-c('r4ss',
            'dplyr',
            'tidyr',
            'here',
            'ggplot2',
            'keyring',
            'DBI',
            'odbc',
            'patchwork',
            'future',
            'stringr',
            'lubridate',
            'ggplot2',                    'afscISS',
            'surveyISS',
            'AreaGrowthRex',
            'spmR',
            'rema')

pacman::p_load(char = newlibs) #load packages
```

## Obtaining GOA rex sole assessment data inputs
#### Open the AFSC database

(assumes username and password are already saved using the keyring package)

```{r}
# Store email username with password
#key_set_with_value(service = "service or email", 
#                   username = "your_address@example.com",
#                   password = "test password")

afsc_user  =  keyring::key_list("afsc")$username  ## enter afsc username
afsc_pwd   = keyring::key_get("afsc", keyring::key_list("afsc")$username)   ## enter afsc password
akfin_user = keyring::key_list("akfin")$username ## enter AKFIN username
akfin_pwd  =  keyring::key_get("akfin", keyring::key_list("akfin")$username)   ## enter AKFIN password


afsc = DBI::dbConnect(odbc::odbc(), "afsc",
                      UID = afsc_user, PWD = afsc_pwd)
akfin = DBI::dbConnect(odbc::odbc(), "akfin",
                       UID = akfin_user, PWD = akfin_pwd)
```

The assessment uses the following data:

### Catch biomass

- One area, as there is only fishing in the western-central GOA and nearly no fishing in the Eastern GOA.

```         
- Download directly from AKFIN Answers ([https://akfinbi.psmfc.org/](https://akfinbi.psmfc.org/bi-security-login/login.jsp?msi=false&redirect=L2FuYWx5dGljcy9zYXcuZGxsP2JpZWVob21lJnN0YXJ0UGFnZT0xJmhhc2g9Y0MtckpJLUVnQUhzUlBtMDM5TWZBamZxdkctT0oydlhlc0VON0h2azlyeFFOSlZRZkNMTVlHYWRNdXRTcmF1Wg==))

  - check boxes for both rex sole and GOA rex sole and grab all years, compare to the previous assessment.
```

### Fishery comp data

####Fishery age comps and sample sizes

- Fishery age data are pulled using the goa_rex case in the sampler repo: <https://github.com/afsc-assessments/sampler/blob/main/cases/rex/call_sampler_rex.R>

  - I used the ADMB code (not the R refactor) for the 2026 assessment

  - I did 1000 bootstraps

  - I could not get sampler to run for the 1992 data, and instead kept the historical data for that year. Although there is 1 haul in that year and it is mostly port data, the same is true for 1995, which does run. The problem would require debugging in the sam.tpl file, as it outputs 0s for age comps.

  - The input sample size for the fishery age data is equal to the number of hauls where ages were collected. There are some years with port data. In 2017 when the fishery age data were incorporated, I found that the historical average number of hauls in a port sample was equal to 9, and this is where the calculation for sample size comes from: it is the number of hauls +9\*(number of port samples), and I calculate this in <https://github.com/afsc-assessments/sampler/blob/main/cases/rex/call_sampler_rex.R>, and it is then merged with the age comp for a final formatted age comp output file catagesex.csv.

#### Fishery length comps

Fishery length composition data are calculated and formatted using the embarrassingly ancient repo newsbss: https://github.com/careymcg/newsbss

The file used is: https://github.com/careymcg/newsbss/blob/master/Inputs_PieceByPiece/Get_Fishery_Lengths/Get_Fishery_Lengths_With_Extrapolated_Number.R

The code pre-dates dplyr!! What a dinosaur! The main piece of code calls functions from this folder: https://github.com/careymcg/newsbss/tree/master/FUNCTIONS

The following specifications are needed and are typically entered in the code pieces mentioned below:

```{r}

DataDir = "C:/Users/carey.mcgilliard/Work/FlatfishAssessments/2026/goa_rex/data/"
dir.create(file.path(DataDir,"Fishery_Lengths"))

len_bins<-seq(from = 9, to = 65, by = 2)

Unsexed = 0 # 1 or 2 (0 = don't include them in comps, 
                     #1 = include unsexed fish in the comps by fraction of m:f in pop
                     #2 = include unsexed fish in the comps as 50% male 50% female


SpeciesCode = "105" #Flathead sole only = 103, Rex = 105, Dover = 107, Bering flounder = 145, GT = 102, Deepsea sole = 110

#GearPerformance: 1 = no problem. however, probably want all gear performances for obs. data for true selex.

#Potentially unbury the choices for FmpArea for max flexibility.
FmpArea = "'GOA'" #Options are 'AI' = 539-544 'GOA' = 600 to 699 'BS' = 500 to 539
#Gear = 1 #For Bering Sea FHS, it's just non-pelagic trawl = 1

# length bins to use for fsh and srv length comp data
bin_width <- 2
min_size <- 9  #9 = Rex #6 = Dover
max_size <- 65
```

There is a separate piece of code for getting the sample sizes: https://github.com/careymcg/newsbss/blob/master/Inputs_PieceByPiece/Get_Fishery_Lengths/Get_Fishery_Length_HaulStats_With_NORPAQ.R

Below are the right specifications for rex sole for the file used for getting the number of hauls for which there are length samples. The code used for getting this info (which is used as sample size for the fishery length comps in the assesmsent input) is:

```{r}
OutDir = "C:\\Users\\carey.mcgilliard\\Work\\FlatfishAssessments\\2026\\goa_rex\\data\\Fishery_Lengths\\"

#
#Flathead sole only = 103, Rex = 105, Dover = 107, Bering flounder = 145, GT = 102, Deepsea sole = 110
#NMFS_AREA between 500 and 544 for BSAI, above 600 for GOA
species = 105
low.nmfs.area = "'600'" #"'600'" #500 
hi.nmfs.area = "'699'" #"'699'"  #544
gear = "all" #"nonpelagic" #"pelagic" #"all"
with_unsexed = F #T
```

Then I read in the output generated by the dinosaur code to format the length comps for ss3, as follows. Note that previous assessments have left out years with low nhauls and I opted to leave these in as they should not impact the objective function much if the number of hauls is very low. I will just list the iss on the plots coming out of r4ss.

```{r}
fishery_lengths<-read.csv(file.path(OutDir,"Observer_Length_Comps.csv")) |>
  dplyr::select(-c(X)) |>
  dplyr::rename(year = YEAR)

fishery_lengths_ss3<-read.csv(file.path(OutDir,"Num_Hauls_ForAndDom_Fishery_AllGears_w_unsexed_FALSE.csv")) |>
  dplyr::select(year,nhauls_tot) |>
  dplyr::right_join(fishery_lengths) |>
  dplyr::rename(nhauls = nhauls_tot) |>
  dplyr::mutate(month = 7, fleet = 1,sex = 3, part = 0) |>
  dplyr::relocate(year, month, fleet, sex, part, nhauls) |>
  tidyr:: drop_na(nhauls) |>
  dplyr::arrange(year) 

write.csv(fishery_lengths_ss3,file = file.path(OutDir,"fishery_lengths_ss3_fmt.csv"))

```

#### Side notes about fishery data

My plan was to update fishery sample size methods when Pete's new repo for processing fishery data is done, which is post-2026 assessment, hence the ancient code that I have not made an effort to update (it seems to still work!)

- *Method used to get fishery ages before 2025 and is no longer used - the method was criticized in a CIE review - it uses raw lengths rather than extrapolated lengths to include the port data): source("C:\\GitProjects\\newsbss\\Inputs_PieceByPiece\\Get_Fishery_Ages\\Get_Fishery_Ages_Port_or_Haul_noStratifying_noExtrapLens.R")*

### Survey Data

- For 2025/2026: see C:/GitProjects/pull_gap_products/examples/goa_rex.qmd which has the code used for pulling all of the survey data and comparing to the 2021 assessment

- Before 2025 (much of this code will not work as it pre-dates gap_products):

  - Survey lengths: source("C:\\GitProjects\\newsbss\\Get_GOA_Survey_Length_and_Age_Comp\\Get_GOA_Length_Comps_by_Area.R")

  - Sample sizes for lengths and ages: source (for nsamp for survey lengths and for ages), by area (made with Wayne Palsson's help) source("C:\\GitProjects\\newsbss\\Get_GOA_Survey_Length_and_Age_Comp\\Get_Survey_Length_Age_Stats_By_Area.R")

  - Note that there are lines for reading in the data from .Rdata files, but below this are the queries for getting the data from SQL and then saving those same .Rdata files #This file also gets the number of hauls. #Don't forget to add in columns for ages that don't appear in the data - this isn't automated. #Run with #FmpArea = "600 and 699" #ByGear = 0 #ByPortOrHaul = 0 #LengthBins = seq(9,65,2) #AgeBins = seq(1,20,1)

  - Survey conditional age-at-lengths #do the SQL query and write the data necessary to read in to the next step here (which will also plot a bunch of age-length ggplots): source("C:\\GitProjects\\newsbss\\Inputs_PieceByPiece\\Get_GOA_Survey_Length_and_Age_Comp\\Get_Survey_Length_Age_and_Plot.R")

  - Run this to get the conditional age-at-length data roughly formatted for assessment input source("C:\\GitProjects\\newsbss\\Inputs_PieceByPiece\\Get_GOA_Survey_Length_and_Age_Comp\\GET_GOA_CONDITIONAL_AGE_AT_LENGTH.R") #run it for "eastern" and for "noneastern" separately #len_bins = seq(9,65,2) #max_age = 20 #eastern = "eastern" then do eastern = "noneastern" #default is 0

  - GOA marginal age comps were tricky because they are not produced by the survey group by area (or by Western-Central and Eastern), otherwise they would be included as ghost comps. (Pete and Ben were trying to resolve this after 2021 using area-specific age-length keys for afscISS and surveyISS. However, I couldn't get their caal data to match mine without any bootstrapping in 2025/2026) \## Configuration notes for ss3 model runs

- the last year of recruitment estimated in the ctl file is typically current year minus 3. In 2026 I set it to current year minus 4 because there was no GOA survey in 2026. Typically rex sole assessments are done in years with GOA data, but this year that is not the case because we never completed the assessment in 2025 due to the October 2025 government shutdown\

- I set the ages included in summary biomass in the starter file to 3 because this is what is reported in the executive summary table. This makes it easier to produce the time series table for total biomass with consistency between ss3 and spm\

- I never bothered to do projections in ss3. I have always used spm and suspect that this model will move to sporc soon, so I didn't do it this year either. A problem for the NPFMC management quantities and ss3 occurs if steepness \<1. Here steepness = 1 (for now), but this could be explored. We have to use SPR-based reference points and biomass reference points use SBPR\*average recruitment, with recruitment as a true average of the estimates, while ss3 assumes that this is average is the mean stock-recruit curve.

## Running the ss3 assessment model and producing model plots

This used to use the newsbss repo and I'm not sure yet if I'm still dependent on this repo, as now I have the AreaGrowthRex package that I made. newsbss used to also be used for data inputs but is not anymore). r4ss is run here: C:/GitProjects/newsbss/Process_SS_Output/SS3toProjectTables.R. Update the fields at the top. This messy code does the following:

- runs r4ss

- makes r4ss plots

- makes additional plots

- calculates updated data weighting

- writes some parameter tables for the document (just simple .csv files)

- writes the data files for doing projections

I am copying in SS3toProjectTables.R to this document to convert it to .qmd

The following relies heavily on r4ss. I did not update r4ss in Aug/Sept 2026, as I heard that functions for doing retrospectives were not working. I repeatedly run the code below, entering various run folder names, then when I pick a base model, I continue past the data weighting chunk to plot things. The file "C:\Users\carey.mcgilliard\Work\FlatfishAssessments\\2026\goa\_rex\runs\GOA\_rex_2026_run_tracker.xlsx" has the run folder names and a description of each run. The way this is left is that the base model is MyRun and SS3PredDir.

### 2026 model run notes

- spr is reported as raw SPR in starter files.\
- Last year of recruitment estimates is current year -3 (consistent with long life in the plankton), but in 2026 we have no survey, the last survey was 2025, so set to current year -4.\
- there were no fishery ages aged yet for 2025\
- The caal for 2025 was tough to get because the ages came the week before the assessment was due and the values for regulatory area used the NMFS definition instead of the INPFC definition used for previous years, and because my re-written code in pull_gap_products did not match that from the special wc and e caal case from surveyISS (and therefore afscISS, which still needs to be worked out)

### Set up plotting themes for manuscript and presentation

```{r plotting themes, include=FALSE}

# Presentation theme
.THEME <- theme(panel.grid.major.x = element_blank(), panel.grid.minor.y = element_blank(), panel.grid.major.y = element_blank() )
.THEME <- .THEME + theme(text=element_text(size=18)) + theme(axis.title.x=element_text(size=24) ,axis.title.y=element_text(size=24, angle = 90,vjust = +2)) +
theme(legend.position = "bottom")
theme_set(.THEME)
#.THEME <- .THEME + theme( panel.background = element_rect(fill="white"), #panel.border = element_rect(colour="black", fill=NA, size=1))

# Manuscript theme
.THEME <- theme(panel.grid.major.x = element_blank(), panel.grid.minor.y = element_blank(), panel.grid.major.y = element_blank() )
.THEME <- .THEME + theme(text=element_text(size=12)) + theme(axis.title.x=element_text(size=12) ,axis.title.y=element_text(size=12, angle = 90,vjust = +2)) +
theme(legend.position = "bottom")
theme_set(.THEME)
```

### Make the single-run r4ss plots for relevant runs

```{r r4ss_plots}

windows(record = T)
par(mfrow = c(1,1),cex.axis = 1.8, cex.lab = 1.8,lwd = 3,omi =c(0.3,0.1,0.3,0.3), mai = c(1.02,1,0.82,0.42),bty = "l")

species = "Rex"#"Halibut" #BSAI_flats #Dover #"Flathead" "Rex" 
run_dir<-"C:/Users/carey.mcgilliard/Work/FlatfishAssessments/2026/goa_rex/runs"
the_run <- "run_19_as_for_18_includes_ageerr"

SS3PredDir = file.path(run_dir,the_run)
setwd(SS3PredDir) #this ends at the end of the chunk

previous_assess_dir<-"C:/Users/carey.mcgilliard/Work/FlatfishAssessments/2021/Rex/Runs"
WorkDir = "C:/Users/carey.mcgilliard/Work/"
CodeDir = "C:/GitProjects/newsbss/"

Uncertainty = T #make r4ss plots with or without uncertainty plotted
FirstYr = 1982 #This is important later for projections and the phase plot.
LastYr = 2026 #LastYr is the last year included in the ss3 model
RecruitmentAge = 3 #RecruitmentAge is the age of recruitment for projections, which are done using the spmR repo
NagesProject = 20 #NagesProject is the plus group age for projections, which is the same as the plus group in ss3

#r4ss, basic plot file, bias adjustment, and Francis lambdas

MyOutput = SS_output(dir = SS3PredDir,covar = Uncertainty,forecast=TRUE,printstats=FALSE)

SS_plots(replist = MyOutput,uncertainty=Uncertainty) 

SSplotPars(replist = MyOutput,showpost = F,fitrange = T, nrow = 8,ncol = 4,pheight= 8,pwidth = 8)
dev.copy(pdf,'plots/SSplotPars.pdf')
dev.off()
```

Make the working directory for saving things SS3PredDir until further notice

```{r change_wd_to_current_run}
#| label: setup
#| include: false
knitr::opts_knit$set(root.dir = SS3PredDir)
```

### Calculate changes in data weighting.

The models (except for a sensitivity analysis) use Francis weighting.

```{r data_weighting}

 Francis <-tune_comps(replist=MyOutput, fleets = "all", option = "Francis",digits = 2,write = TRUE)
# MI<-SS_tune_comps(replist=MyOutput, fleets = "all", option = "MI", digits = 6, write = TRUE)

```

## Jitter analysis

The jitter analysis starts with a jitter fraction of 0.1 and 100 runs. - Convergence Check: If only a tiny fraction (e.g. less than 5 percent) of runs find the lowest likelihood, your model may be highly unstable.\
- Global vs. Local: If many runs find a likelihood lower than your original base model, your original model was stuck in a local minimum. You should adopt the parameter file from the lower likelihood run as your new base. \### Copy the files for jittering

```{r copy_files_for_jittering}
#
# 1. Define directory paths and settings
# Core directory containing your base model files
base_dir<- file.path(run_dir,"run_19_as_for_18_includes_ageerr")

# New directory where jitter runs will be executed
jitter_dir <- file.path(run_dir,"run_19_jitter_runs") 

# 2. Setup the jitter directory
# Copies necessary files (starter, control, data, forecast, ss.exe)
r4ss::copy_SS_inputs(
  dir.old = base_dir,
  dir.new = jitter_dir,
  overwrite = TRUE,
  verbose = TRUE
)
```

### Do the jitter analysis

One note is that the jitter fraction in the starter file in the run folder for jittering must be set to 0.1 (or whatever the jitter_fraction is) manually before running this.

**This section including the call to the jitter() function is commented out so as to not accidentally start 500 jitters by pushing the Render button**

```{r jitter_analysis}
## 3. Execute the jitter loop
# This function automatically modifies starter.ss and runs SS3


# jitter_results <- r4ss::jitter(
#   dir = jitter_dir,            # Directory with model files
#   Njitter = 500,               # Number of random trials (100+ is standard)
#   jitter_fraction = 0.1,        # Your chosen reasonable jitter fraction
#   init_values_src = 0,         # 0 = start each run from the init values in control file
#   printlikes = TRUE,            # Prints likelihoods to console during runs
#   exe = "ss3_win"
# )

## 4. Analyze the results
## Print a summary of the negative log-likelihoods across all runs

#print(jitter_results)

# Count how many runs successfully converged to the global minimum
# (Assuming the lowest value in the vector is the global minimum)
#global_min <- min(jitter_results, na.rm = TRUE)
#success_count <- sum(jitter_results == global_min, na.rm = TRUE)

#cat("Global minimum likelihood found:", global_min, "\n")
#cat("Number of runs that converged to this minimum:", success_count, "out of", #length(jitter_results), "\n")

```

#### Summary of 2026 jitter runs:

Global minimum likelihood found: 793.469\
Number of runs that converged to this minimum: 330 out of 500

This matches the likelihood from the original base case run (run_19) of 793.469, so the .par file was not changed and the model did not need to be initiated with a different parameter vector.

## Make comparison plots

I use and made the R package AreaGrowthRex, originally for a manuscript. It makes nicer comparison plots than r4ss and includes plots of fits to age-length data, as well as maturity-at-age overlaid on selectivity plots.

I updated the 2021 accepted run to use the newest version of ss3 for these plots as the number of columns in the report.sso catch section differs between the version used in 2021 and the version used for this assessment (v3.30.25.1).

### Organize runs for comparison

```{r readfiles, include = FALSE, message = FALSE, warning = FALSE}
compare_dir<-file.path(run_dir,"comparison_plots")
if (!dir.exists(compare_dir)) {
  dir.create(compare_dir, recursive = TRUE)
}
run_2021 <-file.path(run_dir,"run_2021_accepted_new_ss3v")
run_17 <- file.path(run_dir,"run_17_asfor7_newdata_excludes_2025_caal")
run_18 <- file.path(run_dir,"run_18_as_for_17_newdata_includes_2025_caal")
run_base<- file.path(run_dir,"run_19_as_for_18_includes_ageerr")

label_2021<-"2021 accepted"
label_17 <- "new data excludes 2025 caal"
label_18 <- "new data"
label_base <-"base model"

the_run_labels <-c(label_2021,label_17,label_18,label_base)
the_run_dirs<-c(run_2021,run_17,run_18,run_base)

the_run_list<-r4ss::SSgetoutput(dirvec = the_run_dirs,getcomp = TRUE, getcovar = TRUE)

#SummaryList<-r4ss::SSsummarize(MasterList)
```

### Plots comparing time-series estimates

```{r plot estimated pop dynamics, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Figure 6. Observed catches and estimated spawning biomass, recruitment, and fishing intensity (1-spawning potential ratio) for each of the four alternative models with new data added"}


#need to make y-axis on ssb go to zero
ssbplot<-AreaGrowthRex::plot_ssbs(ssruns=the_run_list,mlabel=the_run_labels,showxlabel =   FALSE,showlegend = "none",color_option = "cividis",color_direction =-1)

recplot<-AreaGrowthRex::plot_recs(ssruns = the_run_list,mlabel = the_run_labels,showxlabel = FALSE,showlegend = "none",color_option = "cividis",color_direction = -1)


fi_plot<-AreaGrowthRex::plot_one_minus_spr(ssruns = the_run_list,mlabel = the_run_labels,showxlabel = FALSE,showlegend = "yes",color_option = "cividis",color_direction = -1)

#stuff<-get_catch_data(ssruns = MasterList,mlabel = Mlabels)
total_catch_plot<-AreaGrowthRex::plot_total_catch(ssruns = the_run_list,mlabel = the_run_labels,showxlabel = TRUE,showlegend = "none")

#from library(patchwork)
bigfig<-(total_catch_plot/ssbplot/recplot/fi_plot)

ggplot2::ggsave(filename = file.path(compare_dir,paste0("compare_time_series.png")),plot=bigfig,device = "png",width = 8.5,height = 11)

bigfig
```

### Plots comparing fits to the indices
```{r plot index fits, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Figure 5. Fits (lines) to female and male GOA bottom trawl survey biomass indices for the Western-Central and Eastern GOA."}

p<-AreaGrowthRex::compare_index_fits(ssruns = the_run_list,mlabels=the_run_labels)
p
ggplot2::ggsave(filename = file.path(compare_dir,"index_fits_all_models.png"),device = "png",width = 10, height = 7)
```

### Plots comparing fits to growth data
Specimen data were pulled using pull_gap_products repo and the rex example. Would be nicer if pull_gap_products were an R package and I could use it to re-pull the data needed here.

```{r}
data_dir<-"C:/Users/carey.mcgilliard/Work/FlatfishAssessments/2026/goa_rex/data/output"

specimen_data<-read.csv(file.path(data_dir,"survey_spatial_specimen_data.csv")) %>%
  dplyr::rename_with(toupper)
```

```{r plot growth, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Figure 3. Raw bottom trawl survey data length-at-age by area and sex, overlayed with mean length-at-age for each model (columns) and 90% asymptotic uncertainty intervals around the means (shaded areas). The Eastern GOA is shown in green and the Western-Central GOA is shown in blue."}

g<-AreaGrowthRex::plot_growth(ssruns=the_run_list,sslabels=the_run_labels,thedata= specimen_data, savePlot = FALSE)
g
ggplot2::ggplot2::ggsave(filename = file.path(compare_dir,"Plot_Growth_AllModels.png"),device = "png",width = 10,height = 5)

```

Plot selectivity and maturity
```{r plot selectivity overlaid on maturity, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Figure 7. Fishery and survey selectivity-at-age across models for males and females. Plots for females are overlaid on maturity-at-age"}
p<-AreaGrowthRex::plot_selex_maturity(ssruns=the_run_list,mlabels=the_run_labels) 
ggplot2::ggsave(filename = file.path(compare_dir,paste0("Plot_Selex_Maturity_AllModels.png")),device = "png",width = 10,height = 7)
```

Plot length comps for two-area models. What about the survey lengths from the eastern GOA? Need to work on this plot as these are not included.
```{r plot length comps two area models, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Figure 4. Fits (lines) to female and male fishery length composition with data aggregated by year (shaded regions) for the two-area area model."}
q<-AreaGrowthRex::plot_comps_twoarea(ssruns = the_run_list, narea = 2, mnames = the_run_labels,comptype = "length")
q$combo
ggplot2::ggsave(filename = file.path(compare_dir,paste0("Agg_fits_length_comps_all_models.png")),q$combo,device = "png",width = 10,height = 7)
```
```{r plot fishery ages two area models, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Figure 5. Fits (lines) to female and male fishery age composition with data aggregated by year (shaded regions) for the two-area area model with fishery ages included in model inputs."}

qq<-AreaGrowthRex::plot_comps_twoarea(ssruns = the_run_list, narea = 2, mnames = the_run_labels,comptype = "age")

qq$combo

ggplot2::ggsave(filename = file.path(compare_dir,paste0("Agg_fits_age_comps_all_models.png")),plot = qq$combo,device = "png",width = 10,height = 7)
```
### Some tables
```{r parameter_tables}
#get parameter tables:
compare_params<-AreaGrowthRex::compare_bio_pars(the_run_list,the_run_labels)
write.csv(compare_params,file.path(compare_dir,"compare_time_inv_params.csv"))


```

### Run retrospectives

doRetros is deprecated and there was a bug in 3.30.01. The code below is used but is commented out so that a long process does not start by pressing Render.

```{r run_retrospectives}

# # New directory where jitter runs will be executed
 NRetro = 10
 retro_dir <- file.path(run_dir,"run_19_retro_runs") 
# 
# # 2. Setup the jitter directory
# # Copies necessary files (starter, control, data, forecast, ss.exe)
# r4ss::copy_SS_inputs(
#   dir.old = base_dir,
#   dir.new = retro_dir,
#   overwrite = TRUE,
#   verbose = TRUE
# )
# 
# future::plan(strategy = "multicore")
# retros<-retro(dir = retro_dir, years = 0:-NRetro, exe = "ss3_win")

```
### Plot retrospectives

```{r plot retrospective population dynamics, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Figure 6. Observed catches and estimated spawning biomass, recruitment, and fishing intensity (1-spawning potential ratio) for the retrospective runs"}
NRetro = 10
the_retro_dirs=c(file.path(retro_dir,"retrospectives",paste("retro",1:NRetro,sep="-")),file.path(retro_dir,"retrospectives","retro0"))
the_retro_labels <-c(paste0("retro ",-1:-NRetro),"base model")
the_retro_list <- r4ss::SSgetoutput(dirvec=the_retro_dirs,getcovar = F,forecast = F)
retroSummary <- r4ss::SSsummarize(retroModels)
endyrvec <- retroSummary$endyrs - 0:NRetro

bottom_left_dir<-file.path(compare_dir,"bottom_left")
if (!dir.exists(bottom_left_dir)) {
  dir.create(bottom_left_dir, recursive = TRUE)
}

top_right_dir<-file.path(compare_dir,"top_right")
if (!dir.exists(top_right_dir)) {
  dir.create(top_right_dir, recursive = TRUE)
}

r4ss::SSplotComparisons(retroSummary, plotdir = bottom_left_dir,print = TRUE,endyrvec=endyrvec,xlim = c(1982,2026),legendloc = "bottomleft", legendlabels=paste("Data",0:NRetro,"years"))
dev.off()

r4ss::SSplotComparisons(retroSummary, plotdir = top_right_dir,print = TRUE,endyrvec=endyrvec,xlim = c(1982,2026),legendloc = "topright", legendlabels=paste("Data",0:NRetro,"years"))

write.csv(retroSummary$pars,file = file.path(compare_dir,"RetroParComparison.csv"))
#Calculate Mohn's rho:
#AFSC Mohn's rho:

mohn<-r4ss::SSmohnsrho(retroSummary, endyrvec, verbose = TRUE)

mohn_df<-tibble(ssb = mohn$AFSC_Hurtado_SSB, recruitment = mohn$AFSC_Hurtado_Rec, fishing_mortality = mohn$AFSC_Hurtado_F, bratio = mohn$AFSC_Hurtado_Bratio)

write.csv(mohn_df,file = file.path(compare_dir,"afsc_hurtado_mohn.csv"))
```

### Do OSA residuals


### Do the projections

#### Do the catch estimation
Make sure the akfin connection is open. What is the right package to house the functions used in this chunk? Is it spmR? AreaGrowthRex?  

Note: this does not work for the eastern goa because there are no 2026 catches to go on in the catch estimation function. The egoa catch data are typically confidential and zero or almost zero so filling in spm.dat needs to be done by hand and be careful of the reporting in the SAFE to not report confidential information.  
```{r query detailed catch data}
source("C:/GitProjects/goa_rex/R/query_catch_blend.R", echo = TRUE)
source("C:/GitProjects/goa_rex/R/hp_catch_estimation.R", echo = TRUE)

#query the akfin.council.comprehensive_blend AKRO catch accounting data
wc_catch_data<-query_catch_blend(akfin_conn=akfin,sp_group_code='REXS',region= c('WG','CG'))
e_catch_data<-query_catch_blend(akfin_conn=akfin,sp_group_code='REXS',region= c('WY','EY','SE'))


#write out the latest date for catches queried for reporting (very few catches in egoa so I think using wc for this is ok)
latest_date <-wc_catch_data %>% filter(week_end_date ==max(week_end_date))

write.csv(latest_date,file.path(DataDir,"latest_catch_date_reported.csv"))

# make a wc and an e centralized folder for estimated catch info and past spm.dat
wc_organize_proj_dir <- file.path(DataDir,"wc_proj_info")
e_organize_proj_dir<- file.path(DataDir,"e_proj_info")

  if (!dir.exists(wc_organize_proj_dir)) {
    dir.create(wc_organize_proj_dir, recursive = TRUE)
  }

  if (!dir.exists(e_organize_proj_dir)) {
    dir.create(e_organize_proj_dir, recursive = TRUE)
  }

wc_hp_catch<-hp_catch_estimation(the_dir= wc_organize_proj_dir,catch_data = wc_catch_data,avg_yrs = 5)

#egoa does not have enough data for this query
#e_hp_catch<-hp_catch_estimation(the_dir= wc_organize_proj_dir,catch_data = #e_catch_data,avg_yrs = 5)

e_goa_yearly_catch <-e_catch_data %>% mutate(year= year(week_end_date)) %>%
   group_by(year) %>%
  summarise(sum_catch = sum(weight_posted))
write.csv(e_goa_yearly_catch,file= file.path(e_organize_proj_dir,"e_goa_yearly_catch.csv"))

e_goa_5_yr_avg<-e_goa_yearly_catch %>% 
  filter(year<2026,year>=2026-5) %>%
  ungroup() %>%
  summarize(avg_catch = mean(sum_catch))
write.csv(e_goa_5_yr_avg,file= file.path(e_organize_proj_dir,"e_goa_5yr_avg.csv"))

#Make one centralized place for spm.dat to go and update from the info in the same folder
past_proj_dir<-"C:/Users/carey.mcgilliard/Work/FlatfishAssessments/2024/rex/harvest_projections"

  file.copy(
    from = file.path(past_proj_dir, "western_central","spm.dat"), 
    to = file.path(wc_organize_proj_dir, "spm.dat"), 
    overwrite = FALSE
  )
  
  file.copy(
    from = file.path(past_proj_dir, "eastern","spm.dat"), 
    to = file.path(e_organize_proj_dir, "spm.dat"), 
    overwrite = FALSE
  )  
    
```
#### Transfer realized and estimated catch information to the spm.dat files
Now put the right estimated and realized catch values into the spm.dat file using the wc_organize_proj_dir and e_organize_proj_dir folders created above. Put the catch info in that folder into the spm.dat in the same folder. Do this before the next steps. I could automate this and I haven't.  

##### Update these fields in spm.dat for executive summary table projections:    
beg_yr   
2026 # (end year of the stock assessment model)  
  
data_files for each species                    
ForProjections_GrowthMorph1.dat  #this a western-central example - growth morph 1 is for western-central in the ss3 stuff  
  

fixed_catch              
2026 1 #estimated catch filling in what will be caught at end of year   
2027 1 #5 year average catch  
2028 1 #5 year average catch  
2029 1 #5 year average catch  
  

#### make the .dat files
```{r write_projection_dat_file}

AreaGrowthRex::get_projections(mydir = run_17 ,MyOutput = the_run_list[[2]],NagesProject = 20,FirstYr = 1982,LastYr = 2021,RecruitmentAge=3,species="Rex")

AreaGrowthRex::get_projections(mydir = run_18 ,MyOutput = the_run_list[[3]],NagesProject = 20,FirstYr = 1982,LastYr = 2021,RecruitmentAge=3,species="Rex")

AreaGrowthRex::get_projections(mydir = run_base ,MyOutput = the_run_list[[4]],NagesProject = 20,FirstYr = 1982,LastYr = 2021,RecruitmentAge=3,species="Rex")
```

#### make the folder structure and put files in the right place.
past_proj_dir is where the most recent completed harvest projections are. The spm.dat files should have been updated by hand with realized and estimated catches
```{r}
#define the runs that we want to do projections for
the_proj_dirs<-c(run_18,run_base)
wc_proj_dir <- file.path(the_proj_dirs, "wc_projections")

# Loop through each path in the vector
for (i in seq_along(the_proj_dirs)) {
  # Create directory if it does not exist
  if (!dir.exists(wc_proj_dir[i])) {
    dir.create(wc_proj_dir[i], recursive = TRUE)
  }
  
  # Copy the file to the new directory
  file.copy(
    from = file.path(the_proj_dirs[i], "ForProjections_GrowthMorph1.dat"), 
    to = file.path(wc_proj_dir[i], "ForProjections_GrowthMorph1.dat"), 
    overwrite = TRUE
  )
}

e_proj_dir<-file.path(the_proj_dirs,"e_projections")

# Loop through each path in the vector
for (i in seq_along(the_proj_dirs)) {
  # Create directory if it does not exist
  if (!dir.exists(e_proj_dir[i])) {
    dir.create(e_proj_dir[i], recursive = TRUE)
  }
  
  # Copy the file to the new directory
  file.copy(
    from = file.path(the_proj_dirs[i], "ForProjections_GrowthMorph2.dat"), 
    to = file.path(e_proj_dir[i], "ForProjections_GrowthMorph2.dat"), 
    overwrite = TRUE
  )
  
  #Copy the past hp's tacpar.dat and spm.dat to the new folders
  # Copy the file to the new directory
  file.copy(
    from = file.path(past_proj_dir, "western_central","tacpar.dat"), 
    to = file.path(wc_proj_dir[i], "tacpar.dat"), 
    overwrite = TRUE
  )
    
   file.copy(
    from = file.path(past_proj_dir, "eastern","tacpar.dat"), 
    to = file.path(e_proj_dir[i], "tacpar.dat"), 
    overwrite = TRUE
  )
 
   file.copy(
    from = file.path(wc_organize_proj_dir,"spm.dat"), 
    to = file.path(wc_proj_dir[i], "spm.dat"), 
    overwrite = TRUE
  ) 
    file.copy(
    from = file.path(e_organize_proj_dir,"spm.dat"), 
    to = file.path(e_proj_dir[i], "spm.dat"), 
    overwrite = TRUE
  )
}

```
  
  
**Now check the actual files to make sure everything looks right!**  
  
    
#### Run the projections and make the executive summary table
```{r run_projections_for_list_of_runs}
wc_proj_detail<-list()
e_proj_detail<-list()
for (i in 1:length(wc_proj_dir)) {
wc_proj_detail[[i]] <- spmR::runSPM(wc_proj_dir[i], run = FALSE, engine = "admb")
str(wc_proj_detail)

e_proj_detail[[i]]<- spmR::runSPM(e_proj_dir[i], run = FALSE, engine = "admb")
str(e_proj_detail)
}
```


### Apportionment

Pull the akfin.gap_products.akfin_biomass_v table to get proportion by NMFS regulatory area, including W Yakutat and SE. Would be good to fit the get_spatial_biomass function into a package somewhere for this. There's AreaGrowthRex, rema, pull_gap_products (if it ends up as a package), survey_ISS, and then the goa_rex repo. Where should this go? And the fishery length comp code as well - it should come out of newsbbs and where should it go though?

The views table has a column called nmfs_statistical_area that has the info to split by W Yak and SE outside.
```{r get_spatial_biomass}
source("C:/GitProjects/goa_rex/R/get_spatial_biomass.R", echo = TRUE)
spatial_biomass_data<-get_spatial_biomass(srv_sp_str=c('10200'),surv_id = c('47'),conn = akfin) 

spatial_biomass_data %>% distinct(nmfs_statistical_area)

spatial_biomass_data %>% distinct(regulatory_area) %>%
  filter(regulatory_area == "EASTERN GOA - NMFS") %>%
  distinct(year)

spatial_biomass_data %>% filter(nmfs_statistical_area=="West Yakutat District")

```



#-------------------------------------------------------------------- if (species=="Rex") { #for Rex: plot comparison of maturity curve and selex \# a50\<-5.62298 \# a95\<-10.3917 \# avec\<-3:20

m.df\<-read.csv(file.path(WorkDir,"FlatfishAssessments","2017","Rex","Data","Maturity_At_Age_Buck.csv")) #m.df$est<-1/(1+exp(-log(19)/(a95-a50)*(avec-a50)))
#stock synthesis selectivity:
stuff<-MyOutput$ageselex\[MyOutput$ageselex$Factor == "Asel" & MyOutput$ageselex$Yr == 2021 & MyOutput$ageselex$Sex == 1 & MyOutput$ageselex$Fleet == 1,\] stuff2\<-stuff\[,-c(1:7)\] s.df\<-data.frame(Age = as.numeric(colnames(stuff2)), Sel = as.numeric(stuff2))

plot(0,0,type = "n",bty = "l", xlim = c(0,max(s.df$Age)),ylim = c(0,1),xlab = "Age", ylab = "Proportion")
lines(m.df$Age,m.df$Maturity,lwd = 3, lty = 1, col = "cadetblue")
lines(s.df$Age,s.df\$Sel,lwd = 3, lty = 2, col = "black") legend(x = 13, y = 0.4,legend = c("Maturity","Selectivity (F)"),bty = "n",col = c("cadetblue","black"),lwd = c(3,3), lty = c(1,2))

#save to same folder as for SS plots dev.copy(png,paste0(SS3PredDir,"Maturity_Selex.png")) dev.off() #--------------------------------------------------------------------- source(file.path(CodeDir,"Process_SS_Output","PlotMultipleGrowthMorphs.R")) #save to same folder as for SS plots PlotGrowth(DataDir = file.path(WorkDir,"FlatfishAssessments/2021/Rex/Data/Survey_Age_Length"),AgeLengthData = "Age_Length.csv")

dev.copy(png,file.path(SS3PredDir,"plots","Growth_Curves.png")) dev.off()

} #end if species=="Rex" #------------------------------------------------------------------------------ #check this plotting code - is this plotting dover age-lengths?? if (MyOutput\$ngpatterns == 1) { source(file.path(CodeDir,"Process_SS_Output","PlotSingleGrowthMorphGrowthCurve_And_Data.R")) #this line doesn't work - try C drive instead: dev.copy(png,paste0(redirect,"FlatfishAssessments\\",LastYr,"\\",species,"\\",SS3Folder,"Single_Morph_Growth_Curves.png")) dev.copy(png, "C:\\Data\\Single_Morph_Growth_Curves.png") dev.off() }

#-------------------------------------------------------------------- \# #Tables of param values for a single run (also is with Time_Series_Param_Tables_2_Growth_Morphs.R):

p1.df\<-data.frame(Label = MyOutput$parameters$Label, Value = MyOutput$parameters$Value, Phase = MyOutput$parameters$Phase, StDev = MyOutput$parameters$Parm_StDev)

EstP.df\<-na.omit(p1.df\[p1.df\$Phase \>0,\])

RecPars.df\<-data.frame(type = MyOutput$recruitpars$type, Yr = MyOutput$recruitpars$Yr, Value = MyOutput$recruitpars$Value, StDev = MyOutput$recruitpars$Parm_StDev)

write.csv(RecPars.df,file = paste0(SS3PredDir,"RecPars.csv")) write.csv(EstP.df,file = paste0(SS3PredDir,"Main_Pars.csv")) #easiest with fishing mortality to copy and paste straight from the report file.

#-----------------------------------------------------------------------------

#--------------------------------------------------------------------- #BiasA = SS_fitbiasramp(replist = MyOutput) #Maybe write this to a file? #BiasA #------------------------------------------------------------

#------------------------------------------------------------- #------------------------------------------------------------- #Do projections: #Write out the main projection input file #-------------------------------------------------------------- #before 2017: #source(paste0(redirect,"ProjectionModel\_",LastYr,"\\write_proj_ss_CRM_Dover.r"))

#starting in 2017: #careful: this doesn't work with empirical weight-at-age output; need to manually insert maturity-at-age.

if (species == "Rex" & MyOutput\$ngpatterns \> 1) { #NOTE: NAME ALL DATA FILES STARTING WITH "ForProjections" source(file.path(CodeDir,"Projections","write_proj_ss_CRM_2_growth_morphs.r")) write_proj(data_file=file.path(SS3PredDir,"ForProjections_GrowthMorph1.dat"),data=MyOutput,GrowthMorph=1,area =1,Narea = 2,Nmorph = 2,NAGES=NagesProject,FY=FirstYr,LY=LastYr,RecAge = RecruitmentAge,species = species) write_proj(data_file=file.path(SS3PredDir,"ForProjections_GrowthMorph2.dat"),data=MyOutput,GrowthMorph=2,area =2,Narea = 2,Nmorph = 2,NAGES=NagesProject,FY=FirstYr,LY=LastYr,RecAge = RecruitmentAge,species = species)

} else { #NOTE: NAME ALL DATA FILES STARTING WITH "ForProjections" source(file.path(CodeDir,"Projections","write_proj_ss_CRM.r")) write_proj(data_file=file.path(SS3PredDir,"ForProjections.dat"),data=MyOutput,NAGES=NagesProject,FY=FirstYr,LY=LastYr,RecAge = RecruitmentAge,species = species)

}

#---------------------------------------------------------------------------- #Write a .bat file to run projections within a Projections folder within the run folder #This is handy if running projections for many runs. Otherwise maybe not worth it. #First need to go to FlatfishAssessments/Projections/RexTemplate and update all the files for the current year, including pulling any updates to main.tpl and recompiling. setwd(paste0(SS3PredDir)) data_file = "OrgProjections.bat" #line0\<-noquote(redirect) #write(line0,paste(data_file),ncolumns = 1) line1a\<-noquote(paste0("cd ",'"',SS3PredDir,'"')) write(line1a,paste(data_file),ncolumns = 1) \# myline \<- noquote("mkdir Projections\\data") \# write(myline,paste(data_file),ncolumns = 1, append = T)

if (MyOutput\$ngpatterns ==2) { #NGrowthMorphs \>1 myline \<- noquote("mkdir Projections") write(myline,paste(data_file),ncolumns = 1, append = T)

line2 \<-noquote(paste0("copy ",'"',WorkDir,"\\FlatfishAssessments\\ProjectionModel\\RexTemplate\\GrowthMorph1",'" "',SS3PredDir,"\\Projections\\GrowthMorph1",'"')) line2b \<-noquote(paste0("copy ",'"',WorkDir,"\\FlatfishAssessments\\ProjectionModel\\RexTemplate\\GrowthMorph2",'" "',SS3PredDir,"\\Projections\\GrowthMorph2",'"'))

\# line2b \<-noquote(paste0("copy ",'"',WorkDir,"\\FlatfishAssessments\\ProjectionModel\\RexRunTemplate\\",'" "',SS3PredDir,"Projections\\GrowthMorph2",'"')) write(line2,paste(data_file),ncolumns =1,append = T ) write(line2b,paste(data_file),ncolumns =1,append = T )

myline \<- noquote("mkdir Projections\\GrowthMorph1\\data") write(myline,paste(data_file),ncolumns = 1, append = T) myline2 \<- noquote("mkdir Projections\\GrowthMorph2\\data") write(myline2,paste(data_file),ncolumns = 1, append = T)

line3\<-noquote(paste0("copy ForProjections_GrowthMorph1.dat ",'"',SS3PredDir,"\\Projections\\GrowthMorph1\\data\\ForProjections.dat",'"')) line3b\<-noquote(paste0("copy ForProjections_GrowthMorph2.dat ",'"',SS3PredDir,"\\Projections\\GrowthMorph2\\data\\ForProjections.dat",'"')) write(line3,paste(data_file),ncolumns =1,append = T ) write(line3b,paste(data_file),ncolumns =1,append = T )

line4\<-noquote("cd Projections\\GrowthMorph1\\") write(line4,paste(data_file),ncolumns =1,append = T ) line5\<-noquote("main") write(line5,paste(data_file),ncolumns =1,append = T )

preline\<-noquote("cd..") write(preline,paste(data_file),ncolumns =1,append = T ) line4b\<-noquote("cd GrowthMorph2\\") write(line4b,paste(data_file),ncolumns =1,append = T ) line5b\<-noquote("main") write(line5b,paste(data_file),ncolumns =1,append = T ) }

if (MyOutput\$ngpatterns == 1) { myline \<- noquote("mkdir Projections\\data") write(myline,paste(data_file),ncolumns = 1, append = T) line2 \<-noquote(paste0("copy ",'"',redirect,"\\FlatfishAssessments\\",LastYr,"\\ProjectionModel\\",species,"RunTemplate\\",'" "',SS3PredDir,"Projections",'"'))

write(line2,paste(data_file),ncolumns =1,append = T ) line3\<-noquote(paste0("copy ForProjections\* ",'"',SS3PredDir,"Projections\\data\\ForProjections.dat",'"')) write(line3,paste(data_file),ncolumns =1,append = T ) line4\<-noquote("cd Projections") write(line4,paste(data_file),ncolumns =1,append = T ) line5\<-noquote("main") write(line5,paste(data_file),ncolumns =1,append = T ) } #if NGrowthMorphs == 1 #\*\*\*\*RUN PROJECTIONS\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\*\* shell("OrgProjections")

\`\`\`

## Projections

From ss3toProjecctTables: #CAUTION: for a 2-growth morph model a little help is needed for GP2 projections (GP1 projections are good): #you have to manually copy in the correct .dat file for the projections and change the sppcatch file to the correct GP2 catches

#Before running the projection part of the code below do 2 manual steps: #1. update AK-projection-simulation repo (pull updates) #2. update the projection input files aside from the main (ForProjections) data file (these are in: #C:\Users\carey.mcgilliard\Work\FlatfishAssessments\ProjectionModel\RexTemplate) #Updating requires updated catches and updating the number of years of projections (14 for a full assessment year with a new data) and new catches and check \# years of catches input in proj input files #Probably worthwhile to check into whether it's better to run Jim's R code in AK-projection-simulation repo?? but the code below is for SS output) #Carey can never get Jim's newest projection code to run.

#To: #1. plot SS results #2. Find Francis or McAllister and Ianelli Data weighting update values (for variance adjustments) #3. Plot fishery selex and maturity comparison for rex (as this was what made bio ref pts unreliable in the single area model - fishery selex way to right of maturity) #4. Plot growth data and estimated growth curves for both areas (uses AgeLength.csv, which was made earlier in newsbss/Inputs_PieceByPiece/Get_Survey_Length_Age_and_Plot.R) #5. Write .csv files for making parameter reporting tables #6. Write the projection input file to run folder (Careful: be sure to check over the projection file, especially female and male maturity) #(note writing directly to the run folder is meant to help prevent reporting an exec summary table for the wrong run and also helps when comparing projections for multiple runs) #7. Run projections: source("C:\\GitProjects\\newsbss\\Process_SS_Output\\SS3toProjectTables.R")

#To make the executive summary table and harvest spec tables and do the required phase plot: source(C:\GitProjects\newsbss\Process\_SS_Output\ExecSummaryTable\_Phase_Plot_HarvestRecsTable.R)

#To run retrospectives: #Make a new folder (Like Run_8_Retro), run the following (and this will plot the retros too): #change file paths and run folders source("C:\\GitProjects\\newsbss\\Process_SS_Output\\RunRetrospectives.R")

#Compare any alternative runs using: #change file paths and run folders source("C:\\GitProjects\\newsbss\\Process_SS_Output\\CompareSS3Runs.R")
