---
title: "GOA rex sole assessment instructions"
author: "Carey McGilliard"
format: html
editor: visual
---

# GOA rex sole assessment instructions

## Make some directories to use throughout

```{r make_directories}
#Directories used
past_proj_dir<-"C:/Users/carey.mcgilliard/Work/FlatfishAssessments/2024/rex/harvest_projections"
DataDir = "C:/Users/carey.mcgilliard/Work/FlatfishAssessments/2026/goa_rex/data/"
data_dir<-"C:/Users/carey.mcgilliard/Work/FlatfishAssessments/2026/goa_rex/data/output"
run_dir<-"C:/Users/carey.mcgilliard/Work/FlatfishAssessments/2026/goa_rex/runs"
previous_assess_dir<-"C:/Users/carey.mcgilliard/Work/FlatfishAssessments/2021/Rex/Runs"
WorkDir = "C:/Users/carey.mcgilliard/Work/"
CodeDir = "C:/GitProjects/newsbss/"
OutDir = "C:\\Users\\carey.mcgilliard\\Work\\FlatfishAssessments\\2026\\goa_rex\\data\\Fishery_Lengths\\"
retro_compare_dir<-file.path(run_dir,"retro_comparison_plots")
# make a wc and an e centralized folder for estimated catch info and past spm.dat
wc_organize_proj_dir <- file.path(DataDir,"wc_proj_info")
e_organize_proj_dir<- file.path(DataDir,"e_proj_info")
past_proj_dir<-"C:/Users/carey.mcgilliard/Work/FlatfishAssessments/2024/rex/harvest_projections"

#Organize for comparison plots across runs
run_dir<-"C:/Users/carey.mcgilliard/Work/FlatfishAssessments/2026/goa_rex/runs"
compare_dir<-file.path(run_dir,"comparison_plots")

if (!dir.exists(compare_dir)) {
  dir.create(compare_dir, recursive = TRUE)
}

compare_19_20_dir<-file.path(run_dir,"compare_run_19_run_20")

if (!dir.exists(compare_dir)) {
  dir.create(compare_19_20_dir, recursive = TRUE)
}


end_year= 2026
species = "Rex"#"Halibut" #BSAI_flats #Dover #"Flathead" "Rex" 
the_run <- "run_19_as_for_18_includes_ageerr"

#Runs that we are working with
run_2021 <-file.path(run_dir,"run_2021_accepted_new_ss3v")
#run_17 <- file.path(run_dir,"run_17_asfor7_newdata_excludes_2025_caal")
run_18 <- file.path(run_dir,"run_18_as_for_17_newdata_includes_2025_caal")
run_base<- file.path(run_dir,"run_19_as_for_18_includes_ageerr")
run_20<-file.path(run_dir,"run_20_change_survey_index_timing")
the_run_dirs<-c(run_2021,run_18,run_base)
the_proj_dirs<-c(run_18,run_base)

the_run_dirs_20<-c(run_base, run_20)
the_proj_dirs_20<-c(run_base,run_20)

wc_five_year_proj<-file.path(run_base,"wc_five_year_projections_alt_3")
e_five_year_proj<-file.path(run_base,"e_five_year_projections_alt_3")

if (!dir.exists(wc_five_year_proj)) {
  dir.create(wc_five_year_proj, recursive = TRUE)
}
if (!dir.exists(e_five_year_proj)) {
  dir.create(e_five_year_proj, recursive = TRUE)
}

wc_five_year_proj_ABC<-file.path(run_base,"wc_five_year_projections_ABC")
e_five_year_proj_ABC<-file.path(run_base,"e_five_year_projections_ABC")

if (!dir.exists(wc_five_year_proj)) {
  dir.create(wc_five_year_proj_ABC, recursive = TRUE)
}
if (!dir.exists(e_five_year_proj)) {
  dir.create(e_five_year_proj_ABC, recursive = TRUE)
}

#Labels for the runs
label_2021<-"2021 accepted"
#label_17 <- "new data excludes 2025 caal"
label_18 <- "new data"
label_base <-"base model"
label_20<-"mid-year survey timing"
the_run_labels <-c(label_2021,label_18,label_base)
the_run_labels_20<-c(label_base,label_20)
```

## Read in libraries

```{r}

library(devtools)
devtools::install_github("afsc-assessments/afscISS")

#library(devtools)
#devtools::install_github("BenWilliams-NOAA/surveyISS")

#install repos locally
install.packages("C:/GitProjects/surveyISS", repos = NULL, type = "source")
install.packages("C:/GitProjects/spmR",repos = NULL,type = "source")

devtools::install_github("careymcg/AreaGrowthRex")
devtools::install_github("afsc-assessments/rema", dependencies = TRUE, build_vignettes = TRUE)
# devtools::install_github(repo = "afsc-assessments/spmR", dependencies = TRUE, build_vignettes = TRUE)

if (!requireNamespace("pacman", quietly = TRUE)) {
  install.packages("pacman")
}

newlibs <-c('r4ss',
            'dplyr',
            'tidyr',
            'here',
            'ggplot2',
            'keyring',
            'DBI',
            'odbc',
            'patchwork',
            'future',
            'stringr',
            'lubridate',
            'ggplot2',                    'afscISS',
            'surveyISS',
            'AreaGrowthRex',
            'spmR',
            'rema',
            'knitr')

pacman::p_load(char = newlibs) #load packages
```

## Obtaining GOA rex sole assessment data inputs

#### Open the AFSC database

(assumes username and password are already saved using the keyring package)

```{r open_sql_keyring}

# when your afsc pwd is updated you can easily change it using 
#keyring::key_set_with_value(service="afsc", username="my_username", password = "new_password")

#if there is an extra user on a key:
#key_delete(service = "afsc", username = first_user)
#keyring::key_get("afsc", "MCGILLIARDC")

afsc_user  =  keyring::key_list("afsc")$username  ## enter afsc username
afsc_pwd   = keyring::key_get("afsc", keyring::key_list("afsc")$username)   ## enter afsc password
akfin_user = keyring::key_list("akfin")$username ## enter AKFIN username
akfin_pwd  =  keyring::key_get("akfin", keyring::key_list("akfin")$username)   ## enter AKFIN password


afsc = DBI::dbConnect(odbc::odbc(), "afsc",
                      UID = afsc_user, PWD = afsc_pwd)
akfin = DBI::dbConnect(odbc::odbc(), "akfin",
                       UID = akfin_user, PWD = akfin_pwd)
```

The assessment uses the following data:

### Catch biomass

- One area, as there is only fishing in the western-central GOA and nearly no fishing in the Eastern GOA.

```         
- Download directly from AKFIN Answers ([https://akfinbi.psmfc.org/](https://akfinbi.psmfc.org/bi-security-login/login.jsp?msi=false&redirect=L2FuYWx5dGljcy9zYXcuZGxsP2JpZWVob21lJnN0YXJ0UGFnZT0xJmhhc2g9Y0MtckpJLUVnQUhzUlBtMDM5TWZBamZxdkctT0oydlhlc0VON0h2azlyeFFOSlZRZkNMTVlHYWRNdXRTcmF1Wg==))

  - check boxes for both rex sole and GOA rex sole and grab all years, compare to the previous assessment.
```

### Fishery comp data

####Fishery age comps and sample sizes

- Fishery age data are pulled using the goa_rex case in the sampler repo: <https://github.com/afsc-assessments/sampler/blob/main/cases/rex/call_sampler_rex.R>

  - I used the ADMB code (not the R refactor) for the 2026 assessment

  - I did 1000 bootstraps

  - I could not get sampler to run for the 1992 data, and instead kept the historical data for that year. Although there is 1 haul in that year and it is mostly port data, the same is true for 1995, which does run. The problem would require debugging in the sam.tpl file, as it outputs 0s for age comps.

  - The input sample size for the fishery age data is equal to the number of hauls where ages were collected. There are some years with port data. In 2017 when the fishery age data were incorporated, I found that the historical average number of hauls in a port sample was equal to 9, and this is where the calculation for sample size comes from: it is the number of hauls +9\*(number of port samples), and I calculate this in <https://github.com/afsc-assessments/sampler/blob/main/cases/rex/call_sampler_rex.R>, and it is then merged with the age comp for a final formatted age comp output file catagesex.csv.

#### Fishery length comps

Fishery length composition data are calculated and formatted using the embarrassingly ancient repo newsbss: https://github.com/careymcg/newsbss

The file used is: https://github.com/careymcg/newsbss/blob/master/Inputs_PieceByPiece/Get_Fishery_Lengths/Get_Fishery_Lengths_With_Extrapolated_Number.R

The code pre-dates dplyr!! What a dinosaur! The main piece of code calls functions from this folder: https://github.com/careymcg/newsbss/tree/master/FUNCTIONS

The following specifications are needed and are typically entered in the code pieces mentioned below:

```{r}

dir.create(file.path(DataDir,"Fishery_Lengths"))

len_bins<-seq(from = 9, to = 65, by = 2)

Unsexed = 0 # 1 or 2 (0 = don't include them in comps, 
                     #1 = include unsexed fish in the comps by fraction of m:f in pop
                     #2 = include unsexed fish in the comps as 50% male 50% female


SpeciesCode = "105" #Flathead sole only = 103, Rex = 105, Dover = 107, Bering flounder = 145, GT = 102, Deepsea sole = 110

#GearPerformance: 1 = no problem. however, probably want all gear performances for obs. data for true selex.

#Potentially unbury the choices for FmpArea for max flexibility.
FmpArea = "'GOA'" #Options are 'AI' = 539-544 'GOA' = 600 to 699 'BS' = 500 to 539
#Gear = 1 #For Bering Sea FHS, it's just non-pelagic trawl = 1

# length bins to use for fsh and srv length comp data
bin_width <- 2
min_size <- 9  #9 = Rex #6 = Dover
max_size <- 65
```

There is a separate piece of code for getting the sample sizes: https://github.com/careymcg/newsbss/blob/master/Inputs_PieceByPiece/Get_Fishery_Lengths/Get_Fishery_Length_HaulStats_With_NORPAQ.R

Below are the right specifications for rex sole for the file used for getting the number of hauls for which there are length samples. The code used for getting this info (which is used as sample size for the fishery length comps in the assesmsent input) is:

```{r}

#
#Flathead sole only = 103, Rex = 105, Dover = 107, Bering flounder = 145, GT = 102, Deepsea sole = 110
#NMFS_AREA between 500 and 544 for BSAI, above 600 for GOA
species = 105
low.nmfs.area = "'600'" #"'600'" #500 
hi.nmfs.area = "'699'" #"'699'"  #544
gear = "all" #"nonpelagic" #"pelagic" #"all"
with_unsexed = F #T
```

Then I read in the output generated by the dinosaur code to format the length comps for ss3, as follows. Note that previous assessments have left out years with low nhauls and I opted to leave these in as they should not impact the objective function much if the number of hauls is very low. I will just list the iss on the plots coming out of r4ss.

**Important note: remember to comment out years of fishery length comps for which we have fishery age data so as to not double count the length data - this could be automated in the future**

```{r}
fishery_lengths<-read.csv(file.path(OutDir,"Observer_Length_Comps.csv")) |>
  dplyr::select(-c(X)) |>
  dplyr::rename(year = YEAR)

fishery_lengths_ss3<-read.csv(file.path(OutDir,"Num_Hauls_ForAndDom_Fishery_AllGears_w_unsexed_FALSE.csv")) |>
  dplyr::select(year,nhauls_tot) |>
  dplyr::right_join(fishery_lengths) |>
  dplyr::rename(nhauls = nhauls_tot) |>
  dplyr::mutate(month = 7, fleet = 1,sex = 3, part = 0) |>
  dplyr::relocate(year, month, fleet, sex, part, nhauls) |>
  tidyr:: drop_na(nhauls) |>
  dplyr::arrange(year) 

write.csv(fishery_lengths_ss3,file = file.path(OutDir,"fishery_lengths_ss3_fmt.csv"))
print("time to remove/comment out years of length comps for which there are fishery age data before putting the fishery lengths into the assessment")
```

#### Side notes about fishery data

My plan was to update fishery sample size methods when Pete's new repo for processing fishery data is done, which is post-2026 assessment, hence the ancient code that I have not made an effort to update (it seems to still work!)

- *Method used to get fishery ages before 2025 and is no longer used - the method was criticized in a CIE review - it uses raw lengths rather than extrapolated lengths to include the port data): source("C:\\GitProjects\\newsbss\\Inputs_PieceByPiece\\Get_Fishery_Ages\\Get_Fishery_Ages_Port_or_Haul_noStratifying_noExtrapLens.R")*

### Survey Data

- For 2025/2026: see C:/GitProjects/pull_gap_products/examples/goa_rex.qmd which has the code used for pulling all of the survey data and comparing to the 2021 assessment

- Before 2025 (much of this code will not work as it pre-dates gap_products):

  - Survey lengths: source("C:\\GitProjects\\newsbss\\Get_GOA_Survey_Length_and_Age_Comp\\Get_GOA_Length_Comps_by_Area.R")

  - Sample sizes for lengths and ages: source (for nsamp for survey lengths and for ages), by area (made with Wayne Palsson's help) source("C:\\GitProjects\\newsbss\\Get_GOA_Survey_Length_and_Age_Comp\\Get_Survey_Length_Age_Stats_By_Area.R")

  - Note that there are lines for reading in the data from .Rdata files, but below this are the queries for getting the data from SQL and then saving those same .Rdata files #This file also gets the number of hauls. #Don't forget to add in columns for ages that don't appear in the data - this isn't automated. #Run with #FmpArea = "600 and 699" #ByGear = 0 #ByPortOrHaul = 0 #LengthBins = seq(9,65,2) #AgeBins = seq(1,20,1)

  - Survey conditional age-at-lengths #do the SQL query and write the data necessary to read in to the next step here (which will also plot a bunch of age-length ggplots): source("C:\\GitProjects\\newsbss\\Inputs_PieceByPiece\\Get_GOA_Survey_Length_and_Age_Comp\\Get_Survey_Length_Age_and_Plot.R")

  - Run this to get the conditional age-at-length data roughly formatted for assessment input source("C:\\GitProjects\\newsbss\\Inputs_PieceByPiece\\Get_GOA_Survey_Length_and_Age_Comp\\GET_GOA_CONDITIONAL_AGE_AT_LENGTH.R") #run it for "eastern" and for "noneastern" separately #len_bins = seq(9,65,2) #max_age = 20 #eastern = "eastern" then do eastern = "noneastern" #default is 0

  - GOA marginal age comps were tricky because they are not produced by the survey group by area (or by Western-Central and Eastern), otherwise they would be included as ghost comps. (Pete and Ben were trying to resolve this after 2021 using area-specific age-length keys for afscISS and surveyISS. However, I couldn't get their caal data to match mine without any bootstrapping in 2025/2026) \## Configuration notes for ss3 model runs

- the last year of recruitment estimated in the ctl file is typically current year minus 3. In 2026 I set it to current year minus 4 because there was no GOA survey in 2026. Typically rex sole assessments are done in years with GOA data, but this year that is not the case because we never completed the assessment in 2025 due to the October 2025 government shutdown\

- I set the ages included in summary biomass in the starter file to 3 because this is what is reported in the executive summary table. This makes it easier to produce the time series table for total biomass with consistency between ss3 and spm\

- I never bothered to do projections in ss3. I have always used spm and suspect that this model will move to sporc soon, so I didn't do it this year either. A problem for the NPFMC management quantities and ss3 occurs if steepness \<1. Here steepness = 1 (for now), but this could be explored. We have to use SPR-based reference points and biomass reference points use SBPR\*average recruitment, with recruitment as a true average of the estimates, while ss3 assumes that this is average is the mean stock-recruit curve.

## Running the ss3 assessment model and producing model plots

This used to use the newsbss repo and I'm not sure yet if I'm still dependent on this repo, as now I have the AreaGrowthRex package that I made. newsbss used to also be used for data inputs but is not anymore). r4ss is run here: C:/GitProjects/newsbss/Process_SS_Output/SS3toProjectTables.R. Update the fields at the top. This messy code does the following:

- runs r4ss

- makes r4ss plots

- makes additional plots

- calculates updated data weighting

- writes some parameter tables for the document (just simple .csv files)

- writes the data files for doing projections

I am copying in SS3toProjectTables.R to this document to convert it to .qmd

The following relies heavily on r4ss. I did not update r4ss in Aug/Sept 2026, as I heard that functions for doing retrospectives were not working. I repeatedly run the code below, entering various run folder names, then when I pick a base model, I continue past the data weighting chunk to plot things. The file "C:\Users\carey.mcgilliard\Work\FlatfishAssessments\\2026\goa\_rex\runs\GOA\_rex_2026_run_tracker.xlsx" has the run folder names and a description of each run. The way this is left is that the base model is MyRun and SS3PredDir.

### 2026 model run notes

- spr is reported as raw SPR in starter files.\
- Last year of recruitment estimates is current year -3 (consistent with long life in the plankton), but in 2026 we have no survey, the last survey was 2025, so set to current year -4.\
- there were no fishery ages aged yet for 2025\
- The caal for 2025 was tough to get because the ages came the week before the assessment was due and the values for regulatory area used the NMFS definition instead of the INPFC definition used for previous years, and because my re-written code in pull_gap_products did not match that from the special wc and e caal case from surveyISS (and therefore afscISS, which still needs to be worked out)

### Set up plotting themes for manuscript and presentation

```{r plotting themes, include=FALSE}

# Presentation theme
.THEME <- theme(panel.grid.major.x = element_blank(), panel.grid.minor.y = element_blank(), panel.grid.major.y = element_blank() )
.THEME <- .THEME + theme(text=element_text(size=18)) + theme(axis.title.x=element_text(size=24) ,axis.title.y=element_text(size=24, angle = 90,vjust = +2)) +
theme(legend.position = "bottom")
theme_set(.THEME)
#.THEME <- .THEME + theme( panel.background = element_rect(fill="white"), #panel.border = element_rect(colour="black", fill=NA, size=1))

# Manuscript theme
.THEME <- theme(panel.grid.major.x = element_blank(), panel.grid.minor.y = element_blank(), panel.grid.major.y = element_blank() )
.THEME <- .THEME + theme(text=element_text(size=12)) + theme(axis.title.x=element_text(size=12) ,axis.title.y=element_text(size=12, angle = 90,vjust = +2)) +
theme(legend.position = "bottom")
theme_set(.THEME)
```

### Make the single-run r4ss plots for relevant runs

Don't forget to look at the tunings and update the data weighting in the ss3 input model

```{r r4ss_plots}

#windows(record = T)
par(mfrow = c(1,1),cex.axis = 1.8, cex.lab = 1.8,lwd = 3,omi =c(0.3,0.1,0.3,0.3), mai = c(1.02,1,0.82,0.42),bty = "l")


SS3PredDir = file.path(run_dir,the_run)
setwd(SS3PredDir) #this ends at the end of the chunk


Uncertainty = T #make r4ss plots with or without uncertainty plotted
FirstYr = 1982 #This is important later for projections and the phase plot.
LastYr = 2026 #LastYr is the last year included in the ss3 model
RecruitmentAge = 3 #RecruitmentAge is the age of recruitment for projections, which are done using the spmR repo
NagesProject = 20 #NagesProject is the plus group age for projections, which is the same as the plus group in ss3

#r4ss, basic plot file, bias adjustment, and Francis lambdas

setwd(run_base)
MyOutput = SS_output(dir = run_base,covar = Uncertainty,forecast=TRUE,printstats=FALSE)

#Calculate francis data weighting changes
setwd(run_base)
Francis <-tune_comps(replist=MyOutput, fleets = "all", option = "Francis",digits = 2,write = TRUE)
# MI<-SS_tune_comps(replist=MyOutput, fleets = "all", option = "MI", digits = 6, write = TRUE)

SS_plots(replist = MyOutput,uncertainty=Uncertainty) 

SSplotPars(replist = MyOutput,showpost = F,fitrange = T, nrow = 8,ncol = 4,pheight= 8,pwidth = 8)
dev.copy(pdf,'plots/SSplotPars.pdf')
dev.off()


setwd(run_18)
model_25.0 = SS_output(dir =run_18, covar =T, forecast = TRUE, printstats = FALSE)

SS_plots(replist = model_25.0,uncertainty=Uncertainty) 
SSplotPars(replist = model_25.0,showpost = F,fitrange = T, nrow = 8,ncol = 4,pheight= 8,pwidth = 8)
dev.copy(pdf,'plots/SSplotPars.pdf')


# setwd(run_20)
# model_run_20 = SS_output(dir =run_20, covar =T, forecast = TRUE, printstats = FALSE)
# SS_plots(replist = model_run_20, uncertainty = TRUE)
# SSplotPars(replist = model_run_20,showpost = F,fitrange = T, nrow = 8,ncol = 4,pheight= 8,pwidth = 8)
# dev.copy(pdf,'plots/SSplotPars.pdf')

dev.off()

```
```{r}
source("C:/GitProjects/goa_rex/R/write_proj_ss_CRM_2_growth_morphs_and_SARA.r", echo = TRUE)

write_proj(data_file=file.path(run_base,"ForProjections_GrowthMorph1.dat"),
           sara_file = "Western_Central_Rex_SARA.dat",
           data=MyOutput,GrowthMorph=1,area =1,Narea = 2,Nmorph = 2,NAGES=20,FY=1982,LY=end_year,RecAge = 3, RecAgeSS = 0,species = "REX",Region = "GOA",
                     Tier = "3a",
                     AssessType = "full",
                     SSversion = "SS-V3.30")

write_proj(data_file=file.path(run_base,"ForProjections_GrowthMorph2.dat"),
           sara_file = "Eastern_Rex_SARA.dat",
           data=MyOutput,GrowthMorph=2,area =2,Narea = 2,Nmorph = 2,NAGES=20,FY=1982,LY=end_year,RecAge = 3, RecAgeSS = 0,species = "REX",Region = "GOA",
                     Tier = "3a",
                     AssessType = "full",
                     SSversion = "SS-V3.30")


write_proj(data_file=file.path(run_18,"ForProjections_GrowthMorph1.dat"),
           sara_file = "Western_Central_Rex_SARA.dat",
           data=MyOutput,GrowthMorph=1,area =1,Narea = 2,Nmorph = 2,NAGES=20,FY=1982,LY=end_year,RecAge = 3, RecAgeSS = 0,species = "REX",Region = "GOA",
                     Tier = "3a",
                     AssessType = "full",
                     SSversion = "SS-V3.30")

write_proj(data_file=file.path(run_18,"ForProjections_GrowthMorph2.dat"),
           sara_file = "Eastern_Rex_SARA.dat",
           data=MyOutput,GrowthMorph=2,area =2,Narea = 2,Nmorph = 2,NAGES=20,FY=1982,LY=end_year,RecAge = 3, RecAgeSS = 0,species = "REX",Region = "GOA",
                     Tier = "3a",
                     AssessType = "full",
                     SSversion = "SS-V3.30")

```
Make the working directory for saving things SS3PredDir until further notice

\#`{r change_wd_to_current_run} ##| label: setup ##| include: false #knitr::opts_knit$set(root.dir = SS3PredDir) #`

## Jitter analysis

The jitter analysis starts with a jitter fraction of 0.1 and 100 runs. - Convergence Check: If only a tiny fraction (e.g. less than 5 percent) of runs find the lowest likelihood, your model may be highly unstable.\
- Global vs. Local: If many runs find a likelihood lower than your original base model, your original model was stuck in a local minimum. You should adopt the parameter file from the lower likelihood run as your new base. \### Copy the files for jittering

```{r copy_files_for_jittering}
#
# 1. Define directory paths and settings
# Core directory containing your base model files
base_dir<- file.path(run_dir,"run_19_as_for_18_includes_ageerr")

# New directory where jitter runs will be executed
jitter_dir <- file.path(run_dir,"run_19_jitter_runs") 

# 2. Setup the jitter directory
# Copies necessary files (starter, control, data, forecast, ss.exe)
r4ss::copy_SS_inputs(
  dir.old = base_dir,
  dir.new = jitter_dir,
  overwrite = TRUE,
  verbose = TRUE
)
```

### Do the jitter analysis

One note is that the jitter fraction in the starter file in the run folder for jittering must be set to 0.1 (or whatever the jitter_fraction is) manually before running this.

**This section including the call to the jitter() function is commented out so as to not accidentally start 500 jitters by pushing the Render button**

```{r jitter_analysis}
## 3. Execute the jitter loop
# This function automatically modifies starter.ss and runs SS3


# jitter_results <- r4ss::jitter(
#   dir = jitter_dir,            # Directory with model files
#   Njitter = 500,               # Number of random trials (100+ is standard)
#   jitter_fraction = 0.1,        # Your chosen reasonable jitter fraction
#   init_values_src = 0,         # 0 = start each run from the init values in control file
#   printlikes = TRUE,            # Prints likelihoods to console during runs
#   exe = "ss3_win"
# )

## 4. Analyze the results
## Print a summary of the negative log-likelihoods across all runs

#print(jitter_results)

# Count how many runs successfully converged to the global minimum
# (Assuming the lowest value in the vector is the global minimum)
#global_min <- min(jitter_results, na.rm = TRUE)
#success_count <- sum(jitter_results == global_min, na.rm = TRUE)

#cat("Global minimum likelihood found:", global_min, "\n")
#cat("Number of runs that converged to this minimum:", success_count, "out of", #length(jitter_results), "\n")

```

#### Summary of 2026 jitter runs:

Global minimum likelihood found: 793.469\
Number of runs that converged to this minimum: 330 out of 500

This matches the likelihood from the original base case run (run_19) of 793.469, so the .par file was not changed and the model did not need to be initiated with a different parameter vector.

## Make comparison plots

I use and made the R package AreaGrowthRex, originally for a manuscript. It makes nicer comparison plots than r4ss and includes plots of fits to age-length data, as well as maturity-at-age overlaid on selectivity plots.

I updated the 2021 accepted run to use the newest version of ss3 for these plots as the number of columns in the report.sso catch section differs between the version used in 2021 and the version used for this assessment (v3.30.25.1).

### Organize for comparison plots

```{r organize_comparisons, include = FALSE, message = FALSE, warning = FALSE}


the_run_list<-r4ss::SSgetoutput(dirvec = the_run_dirs,getcomp = TRUE, getcovar = TRUE)
summary_run_list<-r4ss::SSsummarize(the_run_list)
endyrvec <- c(2021,2026, 2026)


bottom_left_dir<-file.path(compare_dir,"bottom_left")
if (!dir.exists(bottom_left_dir)) {
  dir.create(bottom_left_dir, recursive = TRUE)
}

top_right_dir<-file.path(compare_dir,"top_right")
if (!dir.exists(top_right_dir)) {
  dir.create(top_right_dir, recursive = TRUE)
}

r4ss::SSplotComparisons(summary_run_list, plotdir = bottom_left_dir,print = TRUE,endyrvec=endyrvec,xlim = c(1982,2026),legendloc = "bottomleft",legendlabels=the_run_labels)
dev.off()

r4ss::SSplotComparisons(summary_run_list, plotdir = top_right_dir,print = TRUE,endyrvec=endyrvec,xlim = c(1982,2026),legendloc = "topright", legendlabels = the_run_labels)

the_run_labels[1] <-"accepted_model"

write.csv(summary_run_list$pars,file = file.path(compare_dir,"ParComparison.csv"))
```

Very late in the game I discovered that I had survey timing for all of the survey biomass years and areas set to January instead of July. I corrected this in run 20. This came up in the explanation for differences in survey biomass likelihoods between SPoRC and ss3 for Model 25.0 and Model 25.1 Here, I compare run 20 with run 19, the original base case. The results are the same, but not fully identical. The specs will likely be affected in a small way. I am choosing to re-run and add the new results.

```{r organize_comparisons, include = FALSE, message = FALSE, warning = FALSE}
the_run_list_20<-r4ss::SSgetoutput(dirvec = the_run_dirs_20,getcomp = TRUE, getcovar = TRUE)
summary_run_list_20<-r4ss::SSsummarize(the_run_list_20)
endyrvec <- c(2021,2026, 2026)


bottom_left_dir<-file.path(compare_19_20,"bottom_left")
if (!dir.exists(bottom_left_dir)) {
  dir.create(bottom_left_dir, recursive = TRUE)
}

top_right_dir<-file.path(compare_19_20,"top_right")
if (!dir.exists(top_right_dir)) {
  dir.create(top_right_dir, recursive = TRUE)
}

r4ss::SSplotComparisons(summary_run_list_20, plotdir = bottom_left_dir,print = TRUE,endyrvec=endyrvec,xlim = c(1982,2026),legendloc = "bottomleft",legendlabels=the_run_labels_20)
dev.off()

r4ss::SSplotComparisons(summary_run_list_20, plotdir = top_right_dir,print = TRUE,endyrvec=endyrvec,xlim = c(1982,2026),legendloc = "topright", legendlabels = the_run_labels_20)

the_run_labels[1] <-"accepted_model"

write.csv(summary_run_list_20$pars,file = file.path(compare_19_20,"ParComparison.csv"))
```

### Plots comparing time-series estimates

```{r plot estimated pop dynamics, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Figure 6. Observed catches and estimated spawning biomass, recruitment, and fishing intensity (1-spawning potential ratio) for each of the four alternative models with new data added"}

#need ssbplot for the base case for stock assessment guidelines
ssb_base_plot<-AreaGrowthRex::plot_ssbs(ssruns = the_run_list[3], mlabel = the_run_labels[3], showxlabel = TRUE, showlegend = "yes",color_option = "cividis",color_direction =1)
ggplot2::ggsave(filename = file.path(run_base,"plots",paste0("ssb_base_case.png")),plot = ssb_base_plot)

#need to make y-axis on ssb go to zero
ssbplot<-AreaGrowthRex::plot_ssbs(ssruns=the_run_list,mlabel=the_run_labels,showxlabel =   FALSE,showlegend = "none",color_option = "cividis",color_direction =-1)

recplot<-AreaGrowthRex::plot_recs(ssruns = the_run_list,mlabel = the_run_labels,showxlabel = FALSE,showlegend = "none",color_option = "cividis",color_direction = -1)


fi_plot<-AreaGrowthRex::plot_one_minus_spr(ssruns = the_run_list,mlabel = the_run_labels,showxlabel = FALSE,showlegend = "yes",color_option = "cividis",color_direction = -1)

#stuff<-get_catch_data(ssruns = MasterList,mlabel = Mlabels)
total_catch_plot<-AreaGrowthRex::plot_total_catch(ssruns = the_run_list,mlabel = the_run_labels,showxlabel = TRUE,showlegend = "none")

#from library(patchwork)
bigfig<-(total_catch_plot/ssbplot/recplot/fi_plot)

ggplot2::ggsave(filename = file.path(compare_dir,paste0("compare_time_series.png")),plot=bigfig,device = "png",width = 8.5,height = 11)

bigfig
```

### Plots comparing fits to the indices

```{r plot index fits, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Figure 5. Fits (lines) to female and male GOA bottom trawl survey biomass indices for the Western-Central and Eastern GOA."}

p<-AreaGrowthRex::compare_index_fits(ssruns = the_run_list,mlabels=the_run_labels)
p
ggplot2::ggsave(filename = file.path(compare_dir,"index_fits_all_models.png"),device = "png",width = 10, height = 7)
```

### Plots comparing fits to growth data

Specimen data were pulled using pull_gap_products repo and the rex example. Would be nicer if pull_gap_products were an R package and I could use it to re-pull the data needed here.

```{r}

specimen_data<-read.csv(file.path(data_dir,"survey_spatial_specimen_data.csv")) %>%
  dplyr::rename_with(toupper)
```

```{r plot growth, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Figure 3. Raw bottom trawl survey data length-at-age by area and sex, overlayed with mean length-at-age for each model (columns) and 90% asymptotic uncertainty intervals around the means (shaded areas). The Eastern GOA is shown in green and the Western-Central GOA is shown in blue."}

g<-AreaGrowthRex::plot_growth(ssruns=the_run_list,sslabels=the_run_labels,thedata= specimen_data, savePlot = FALSE)
g
ggplot2::ggsave(filename = file.path(compare_dir,"Plot_Growth_AllModels.png"),device = "png",width = 10,height = 5)

```

Plot selectivity and maturity

```{r plot selectivity overlaid on maturity, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Figure 7. Fishery and survey selectivity-at-age across models for males and females. Plots for females are overlaid on maturity-at-age"}
p<-AreaGrowthRex::plot_selex_maturity(ssruns=the_run_list,mlabels=the_run_labels) 
p
ggplot2::ggsave(filename = file.path(compare_dir,paste0("Plot_Selex_Maturity_AllModels.png")),device = "png",width = 10,height = 7)
```

Plot length comps for two-area models. What about the survey lengths from the eastern GOA? Need to work on this plot as these are not included.

```{r plot length comps two area models, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Figure 4. Fits (lines) to female and male fishery length composition with data aggregated by year (shaded regions) for the two-area area model."}
q<-AreaGrowthRex::plot_comps_twoarea(ssruns = the_run_list, narea = 2, mnames = the_run_labels,comptype = "length")
q$combo
ggplot2::ggsave(filename = file.path(compare_dir,paste0("Agg_fits_length_comps_all_models.png")),q$combo,device = "png",width = 10,height = 7)

```

```{r plot fishery ages two area models, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Figure 5. Fits (lines) to female and male fishery age composition with data aggregated by year (shaded regions) for the two-area area model with fishery ages included in model inputs."}

qq<-AreaGrowthRex::plot_comps_twoarea(ssruns = the_run_list, narea = 2, mnames = the_run_labels,comptype = "age")

qq$combo

ggplot2::ggsave(filename = file.path(compare_dir,paste0("Agg_fits_age_comps_all_models.png")),plot = qq$combo,device = "png",width = 10,height = 7)
```

### Some tables

```{r parameter_tables}
#get parameter tables:
compare_params<-AreaGrowthRex::compare_bio_pars(the_run_list,the_run_labels)
write.csv(compare_params,file.path(compare_dir,"compare_time_inv_params.csv"))
```

### Run retrospectives

doRetros is deprecated and there was a bug in 3.30.01. The code below is used but is commented out so that a long process does not start by pressing Render.

```{r run_retrospectives}

# # New directory where jitter runs will be executed
 NRetro = 10
 retro_dir <- file.path(run_dir,"run_19_retro_runs") 

# 2. Setup the jitter directory
# Copies necessary files (starter, control, data, forecast, ss.exe)
r4ss::copy_SS_inputs(
  dir.old = run_base,
  dir.new = retro_dir,
  overwrite = TRUE,
  verbose = TRUE
)

future::plan(strategy = "multicore")
retros<-retro(dir = retro_dir, years = 0:-NRetro, exe = "ss3_win")

```

### Plot retrospectives

```{r plot retrospective population dynamics, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Figure 6. Observed catches and estimated spawning biomass, recruitment, and fishing intensity (1-spawning potential ratio) for the retrospective runs"}
NRetro = 10
the_retro_dirs=c(file.path(retro_dir,"retrospectives",paste("retro",1:NRetro,sep="-")),file.path(retro_dir,"retrospectives","retro0"))
the_retro_labels <-c(paste0("retro ",-1:-NRetro),"base model")
the_retro_list <- r4ss::SSgetoutput(dirvec=the_retro_dirs,getcovar = F,forecast = F)
retroSummary <- r4ss::SSsummarize(the_retro_list)
endyrvec <- retroSummary$endyrs - 0:NRetro

bottom_left_dir<-file.path(retro_compare_dir,"bottom_left")
if (!dir.exists(bottom_left_dir)) {
  dir.create(bottom_left_dir, recursive = TRUE)
}

top_right_dir<-file.path(retro_compare_dir,"top_right")
if (!dir.exists(top_right_dir)) {
  dir.create(top_right_dir, recursive = TRUE)
}

r4ss::SSplotComparisons(retroSummary, plotdir = bottom_left_dir,print = TRUE,endyrvec=endyrvec,xlim = c(1982,2026),legendloc = "bottomleft", legendlabels=paste("Data",0:NRetro,"years"))
dev.off()

r4ss::SSplotComparisons(retroSummary, plotdir = top_right_dir,print = TRUE,endyrvec=endyrvec,xlim = c(1982,2026),legendloc = "topright", legendlabels=paste("Data",0:NRetro,"years"))

write.csv(retroSummary$pars,file = file.path(retro_compare_dir,"RetroParComparison.csv"))
#Calculate Mohn's rho:
#AFSC Mohn's rho:

mohn<-r4ss::SSmohnsrho(retroSummary, endyrvec, verbose = TRUE)

mohn_df<-tibble(ssb = mohn$AFSC_Hurtado_SSB, recruitment = mohn$AFSC_Hurtado_Rec, fishing_mortality = mohn$AFSC_Hurtado_F, bratio = mohn$AFSC_Hurtado_Bratio)

write.csv(mohn_df,file = file.path(compare_dir,"afsc_hurtado_mohn.csv"))
```

### Phase plot

### Do OSA residuals

### Do the projections

#### Catch estimation

Make sure the akfin connection is open. What is the right package to house the functions used in this chunk? Is it spmR? AreaGrowthRex?

Note: this does not work for the eastern goa because there are no 2026 catches to go on in the catch estimation function. The egoa catch data are typically confidential and zero or almost zero so filling in spm.dat needs to be done by hand and be careful of the reporting in the SAFE to not report confidential information.

```{r query detailed catch data}
end_year = 2026

source("C:/GitProjects/goa_rex/R/query_catch_blend.R", echo = TRUE)
source("C:/GitProjects/goa_rex/R/hp_catch_estimation.R", echo = TRUE)

#query the akfin.council.comprehensive_blend AKRO catch accounting data
wc_catch_data<-query_catch_blend(akfin_conn=akfin,sp_group_code='REXS',region= c('WG','CG'))
e_catch_data<-query_catch_blend(akfin_conn=akfin,sp_group_code='REXS',region= c('WY','EY','SE'))


#write out the latest date for catches queried for reporting (very few catches in egoa so I think using wc for this is ok)
latest_date <-wc_catch_data %>% filter(week_end_date ==max(week_end_date))

write.csv(latest_date,file.path(DataDir,"latest_catch_date_reported.csv"))

# make a wc and an e centralized folder for estimated catch info and past spm.dat
#wc_organize_proj_dir <- file.path(DataDir,"wc_proj_info")
#e_organize_proj_dir<- file.path(DataDir,"e_proj_info")

  if (!dir.exists(wc_organize_proj_dir)) {
    dir.create(wc_organize_proj_dir, recursive = TRUE)
  }

  if (!dir.exists(e_organize_proj_dir)) {
    dir.create(e_organize_proj_dir, recursive = TRUE)
  }

wc_hp_catch<-hp_catch_estimation(the_dir= wc_organize_proj_dir,catch_data = wc_catch_data,avg_yrs = 5)

#egoa does not have enough data for this query
#e_hp_catch<-hp_catch_estimation(the_dir= wc_organize_proj_dir,catch_data = #e_catch_data,avg_yrs = 5)

e_goa_yearly_catch <-e_catch_data %>% mutate(year= year(week_end_date)) %>%
   group_by(year) %>%
  summarise(sum_catch = sum(weight_posted))
write.csv(e_goa_yearly_catch,file= file.path(e_organize_proj_dir,"e_goa_yearly_catch.csv"))

e_goa_5_yr_avg<-e_goa_yearly_catch %>% 
  filter(year<end_year,year>=end_year-5) %>%
  ungroup() %>%
  summarize(avg_catch = mean(sum_catch))
write.csv(e_goa_5_yr_avg,file= file.path(e_organize_proj_dir,"e_goa_5yr_avg.csv"))

#Make one centralized place for spm.dat to go and update from the info in the same folder
#past_proj_dir<-"C:/Users/carey.mcgilliard/Work/FlatfishAssessments/2024/rex/harvest_projections"

  file.copy(
    from = file.path(past_proj_dir, "western_central","spm.dat"), 
    to = file.path(wc_organize_proj_dir, "spm.dat"), 
    overwrite = FALSE
  )
  
  file.copy(
    from = file.path(past_proj_dir, "eastern","spm.dat"), 
    to = file.path(e_organize_proj_dir, "spm.dat"), 
    overwrite = FALSE
  )  
    
```

#### Transfer realized and estimated catch information to the spm.dat files

Now put the right estimated and realized catch values into the spm.dat file using the wc_organize_proj_dir and e_organize_proj_dir folders created above. Put the catch info in that folder into the spm.dat in the same folder. Do this before the next steps. I could automate this and I haven't.

##### Update these fields in spm.dat for executive summary table projections:

beg_yr\
2026 \# (end year of the stock assessment model)

data_files for each species\
ForProjections_GrowthMorph1.dat #this a western-central example - growth morph 1 is for western-central in the ss3 stuff

fixed_catch\
2026 1450 #estimated catch filling in what will be caught at end of year\
2027 1700 #5 year average catch\
2028 1700 #5 year average catch\
2029 1700 #5 year average catch

#### make the .dat files

I think there is a problem here where spm.tpl was changed and something (recruitment?) might need to be doubled when making these .dat files. I will check into it tonight.

```{r write_projection_dat_file}

AreaGrowthRex::get_projections(mydir = run_2021 ,MyOutput = the_run_list[[1]],NagesProject = 20,FirstYr = 1982,LastYr = 2021,RecruitmentAge=3,species="Rex")

AreaGrowthRex::get_projections(mydir = run_18 ,MyOutput = the_run_list[[2]],NagesProject = 20,FirstYr = 1982,LastYr = end_year,RecruitmentAge=3,species="Rex")

AreaGrowthRex::get_projections(mydir = run_base ,MyOutput = the_run_list[[3]],NagesProject = 20,FirstYr = 1982,LastYr = end_year,RecruitmentAge=3,species="Rex")
```

#### make the folder structure and put files in the right place.

past_proj_dir is where the most recent completed harvest projections are. The spm.dat files should have been updated by hand with realized and estimated catches. This chunk should mostly be a function with the_proj_dirs as input.

*Don't run the next chunk for the 2021 model - don't overwrite the spm.dat file for the last accepted model, OR go back and re-copy in the previous spm.dat to the new folder run before making the executive summary table.*

Most of the run directories are specified in the chunk @organize_comparisons

```{r}
#| label: organize_projections


#the_proj_dirs<-c(base_old_ss3_dir) #for troubleshooting, start below
#the_proj_dirs<-(c(age0_smry_bio_dir)) #for troubleshooting, start below
#the_proj_dirs<-(c(run_2021_new_ss3_dir))
#the_proj_dirs<-(c(run_19_old_spm_dir))
#the_proj_dirs<-(c(run_19_half_rec_dir))

wc_proj_dir <- file.path(the_proj_dirs, "wc_projections")

# Loop through each path in the vector
for (i in seq_along(the_proj_dirs)) {
  # Create directory if it does not exist
  if (!dir.exists(wc_proj_dir[i])) {
    dir.create(wc_proj_dir[i], recursive = TRUE)
  }
  
  # Copy the file to the new directory
  file.copy(
    from = file.path(the_proj_dirs[i], "ForProjections_GrowthMorph1.dat"), 
    to = file.path(wc_proj_dir[i], "ForProjections_GrowthMorph1.dat"), 
    overwrite = TRUE
  )
}

e_proj_dir<-file.path(the_proj_dirs,"e_projections")

# Loop through each path in the vector
for (i in seq_along(the_proj_dirs)) {
  # Create directory if it does not exist
  if (!dir.exists(e_proj_dir[i])) {
    dir.create(e_proj_dir[i], recursive = TRUE)
  }
  
  # Copy the file to the new directory
  file.copy(
    from = file.path(the_proj_dirs[i], "ForProjections_GrowthMorph2.dat"), 
    to = file.path(e_proj_dir[i], "ForProjections_GrowthMorph2.dat"), 
    overwrite = TRUE
  )
  
  #Copy the past hp's tacpar.dat and spm.dat to the new folders
  # Copy the file to the new directory
  file.copy(
    from = file.path(past_proj_dir, "western_central","tacpar.dat"), 
    to = file.path(wc_proj_dir[i], "tacpar.dat"), 
    overwrite = TRUE
  )
    
   file.copy(
    from = file.path(past_proj_dir, "eastern","tacpar.dat"), 
    to = file.path(e_proj_dir[i], "tacpar.dat"), 
    overwrite = TRUE
  )
 
   file.copy(
    from = file.path(wc_organize_proj_dir,"spm.dat"), 
    to = file.path(wc_proj_dir[i], "spm.dat"), 
    overwrite = TRUE
  ) 
    file.copy(
    from = file.path(e_organize_proj_dir,"spm.dat"), 
    to = file.path(e_proj_dir[i], "spm.dat"), 
    overwrite = TRUE
  )
}

```

**Now check the actual files to make sure everything looks right!**

#### Run the projections and make the executive summary table

\*Important note: In spm.tpl, there was a change between 2021 and 2024 where previously we input total recruitment across sexes and in the current version we need to input recruitment for only 1 sex, or total recruitment/2.

Lines 630-634 in the PARAMETER_SECTION under LOCAL_CALCS (and you can see that the earlier assumption is commented out, and that this applies to 2-sex model output. The solution is to fix this in the AreaGrowthRex package under R/write_proj_ss_2Area.R and R/write_proj_ss_1Area.R by either only entering female recruits or by dividing in half prior to writing it out.\*

````         
```         
if (nsexes(ispp)==1) 
  AMeanRec(ispp) = mean(R(ispp));  // Arithmetic mean
else
  AMeanRec(ispp) = 2.*mean(R(ispp));  // Arithmetic mean, converted to females (half) later
  // AMeanRec(ispp) = mean(R(ispp));  // Arithmetic mean, converted to females (half) later
```
````

spmR works best by cloning the repo and compiling spm.tpl yourself, putting it in your path or projection folder and running the .exe, then using the spmR R functions. The following will work after running spm in each projection folder. I ran the projections manually and I copied in spm.exe, which I first compiled with ADMB 13.0.The rtmb implementation didnt work either. I also don't know why the make_tier3a_exec_table.R table is not integrated into the spmR package and need to get it in there after assessments.

```{r}
#| label: make_exec_summary_tables
 #source("C:/GitProjects/spmR/R/make_tier3a_exec_table.R", echo = TRUE)
source("C:/GitProjects/spmR/R/make_tier3a_exec_table.R", echo = TRUE)

wc_proj_detail<-list()
e_proj_detail<-list()
wc_exec<-list()
e_exec<-list()
end_year_vec = c(2026,2026) # in case running new projections for an older year model

for (i in 1:length(wc_proj_dir)) {
wc_proj_detail[[i]] <- spmR::runSPM(wc_proj_dir[i], run = FALSE, engine = "admb")
str(wc_proj_detail)

wc_exec[i]<-make_tier3a_exec_table(wc_proj_dir[i],endyr=end_year_vec[i],the_scalar=1000)

e_proj_detail[[i]]<- spmR::runSPM(e_proj_dir[i], run = FALSE, engine = "admb")
str(e_proj_detail)

e_exec[i]<-make_tier3a_exec_table(e_proj_dir[i],endyr=end_year_vec[i],the_scalar=1000)
}

```

Make the scenario projection tables

```{r make_scenario_projection tables}

proj_catches_wc<-scenario_tables(proj_dir = wc_proj_dir[[2]],
                                 spm_detail_df = wc_proj_detail[[i]],the_scalar = 1000)

proj_catches_e<-scenario_tables(proj_dir = e_proj_dir[[2]],
                                 spm_detail_df = e_proj_detail[[i]],the_scalar = 1000)
```

### 5 year projections

For rex sole because it's a spatial model we have to run the five year projection table after doing apportionment because we need western+central ABC and eastern combined abc for years 2027 and 2028.

Here are the 2026 instructions for this: Provide a table of 5 year projections using the author recommend ABC for model end years +1 and +2 based on the author’s catch projection recommendation, and Scenario 3, as described in the Standard Harvest Scenarios and Projection Methodology, for model end years +3 to +5 with the following headers: Year, Expected Catch, OFL, ABC, Mean SSB, Mean Relative Spawning Biomass (ratio of female spawning biomass to equilibrium unfished female spawning biomass (B100%)).

Scenario 3 is: In all future years, F is set equal to 50% of max FABC. (Rationale: This scenario provides a likely lower bound on FABC that still allows future harvest rates to be adjusted downward when stocks fall below reference levels.)

*Note: to do scenario 3 and write out the correct ABC, you have to use MaxABC, which is now incorporated into the five_year_hp_table function. Pull Alt 3 results. Run spm with the recommended ABCs from the executive summary table for 2027 and 2028, and unspecified 2029-2031 catches.*

The projection model runs for 5 year table: Alt 3

```{r five_year_projections_part_1}
source("C:/GitProjects/goa_rex/R/five_year_hp_table.R", echo = TRUE)

#these are the 5-year projection tables, except that ABC is wrong.
wc_five_year_alt_3_df<-five_year_hp_table(hp_dir =wc_five_year_proj ,current_year = year(today()),alt = 3)

e_five_year_alt_3_df<-five_year_hp_table(hp_dir =e_five_year_proj ,current_year = year(today()),alt = 3) # %>%  dplyr:: mutate(mean_catch = as.character(mean_catch),mean_catch = if_else(Year > 2028, "C", mean_catch))

```

<!-- The second projection model run to get the right ABCs for the 5-year projection table. -->

<!-- Now take the catches that come from running with Alt = 3 and put them in spm.dat (manually) and re-run spm in a new folder (manually). Then re-make the five-year projection tables using Alt 1 to get the right ABC that goes along with catches that come from the fixed F (alt 3) run. -->

<!-- ```{r five_year_projections_final} -->
<!-- source("C:/GitProjects/goa_rex/R/five_year_hp_table.R", echo = TRUE) -->

<!-- #these are the 5-year projection tables, except that ABC is wrong. -->
<!-- wc_five_year_ABC_df<-five_year_hp_table(hp_dir =wc_five_year_proj_ABC ,current_year = year(today()),alt = 1) -->

<!-- e_five_year_ABC_df<-five_year_hp_table(hp_dir =e_five_year_proj_ABC ,current_year = year(today()),alt = 1) # %>%  dplyr:: mutate(mean_catch = as.character(mean_catch),mean_catch = if_else(Year > 2028, "C", mean_catch)) -->


<!-- ``` -->

### Apportionment

Pull the akfin.gap_products.akfin_biomass_v table to get proportion by NMFS regulatory area, including W Yakutat and SE. Would be good to fit the get_spatial_biomass function into a package somewhere for this. There's AreaGrowthRex, rema, pull_gap_products (if it ends up as a package), survey_ISS, and then the goa_rex repo. Where should this go? And the fishery length comp code as well - it should come out of newsbbs and where should it go though?

The views table has a column called nmfs_statistical_area that has the info to split by W Yak and SE outside.

split_fractions is a table by year listing the proportion of biomass in WYak vs SE for years when survey aligned with INPFC areas, which is all years before 2025. WYak and SE are NMFS areas, so this was a bespoke method. Now the survey aligns with NMFS statistical areas, so the spatial biomass data in akfin.gap_products.akfin_biomass_v using the column nmfs_statistical_area should align from 2025 onward. We will always need to cobble together the full biomass series from these two periods: pre-2025 and onward.

```{r egoa_biomass_by_stat_area}
source("C:/GitProjects/goa_rex/R/get_biomass_by_stat_area.R", echo = TRUE)
biomass_rema_df<-get_biomass_by_stat_area(conn = akfin,data_dir=data_dir,sp_group_code = 'rex sole',survey_species_code=c('10200'),surv_id = c('47'))

#table for the doc
spatial_biomass_df<-biomass_rema_df %>%
  tidyr::pivot_wider(id_cols = year,names_from = strata,values_from = c(biomass,cv)) %>%
  dplyr::relocate(biomass_wgoa,cv_wgoa,biomass_cgoa,cv_cgoa,biomass_wyak,cv_wyak,biomass_se,cv_se)

write.csv(spatial_biomass_df,file.path(DataDir,"spatial_survey_biomass_table.csv"),na = "")
```

Set up and run rema, first with an assumption of shared process error over strata (areas). this doesn't work with the following error message: Error in rema::prepare_rema_input(model_name = "rema goa rex sole", biomass_dat = biomass_rema_df, : user defined multi_survey as TRUE but did not provide CPUE survey data in the admb_re list or as a dataframe in the cpue_dat argument.

```{r run_rema_shared_pe}

# the original ADMB model shared PE across all strata
input_shared_pe <- rema::prepare_rema_input(model_name = 'rema goa rex sole shared pe',
                            biomass_dat = biomass_rema_df,
                            multi_survey = 1,
                            PE_options = list(pointer_PE_biomass = c(1, 1, 1)))
model_shared_pe <- fit_rema(input_shared_pe)
output_shared_pe <- tidy_rema(rema_model = model_shared_pe)
kable(output_shared_pe$parameter_estimates) # estimated fixed effects parameters
```

Next, run without the assumption of shared process error. This works, but for rex I need to run one model for western and central data and another for Wyak and SE.

```{r run_rema}
wc_biomass_rema_df<-biomass_rema_df %>% filter(strata == 'cgoa' | strata == 'wgoa')
wc_input <- rema::prepare_rema_input(model_name = 'wc rema goa rex sole',
                            biomass_dat = wc_biomass_rema_df,
                            multi_survey = 0)
wc_model <- fit_rema(wc_input)
wc_output<- tidy_rema(rema_model = wc_model)
wc_est_params<-wc_output$parameter_estimates # estimated fixed effects parameters
wc_area_proportions<-wc_output$proportion_biomass_by_strata %>% filter(year==2025)
write.csv(wc_output$proportion_biomass_by_strata,file = file.path(data_dir,"wc_rema_proportions.csv"))
```

Eastern GOA rema model

```{r run_rema_egoa}
e_biomass_rema_df<-biomass_rema_df %>% filter(strata == 'wyak' | strata == 'se')
e_input <- rema::prepare_rema_input(model_name = 'e rema goa rex sole',
                            biomass_dat = e_biomass_rema_df,
                            multi_survey = 0)
e_model <- fit_rema(e_input)
e_output<- tidy_rema(rema_model = e_model)
e_est_params<-e_output$parameter_estimates # estimated fixed effects parameters
e_area_proportions<-e_output$proportion_biomass_by_strata %>% filter(year==2025)
write.csv(e_output$proportion_biomass_by_strata,file = file.path(data_dir,"e_rema_proportions.csv"))
```

Long projection tables

```{r}


```

## Making survey cpue maps

Matt Callahan made these for me. They come from this code:

Kalei says Matt used to make IDW maps (inverse weighted distance), but the survey group mentioned that they are innappropriate for GOA/AI survey data and for much marine ecological data. Kalei asked gemini why that would be a bad method and this is what it said:\
"Yes, using Inverse Distance Weighting (IDW) to interpolate Catch Per Unit Effort (CPUE) haul data, as seen in "Screenshot 2026-08-27 161053.png" for shortraker rockfish in the Aleutian Islands, is generally considered inappropriate for marine ecological data.\
While IDW is easy to compute and visually striking, its underlying mathematical assumptions conflict with the biological realities of how marine species are distributed. Here are the primary reasons why this method falls short for this type of fisheries data:\
**Ignores Depth and Habitat Boundaries:** IDW calculates values based purely on 2D geographic distance. It assumes that if you have a high catch at Point A, a location physically close to it (Point B) should also have a high catch. However, marine environments—especially in the Aleutian Islands—feature steep drop-offs and complex bathymetry. Point A might be at an optimal depth of 300 meters, while Point B, just a mile away, might be in a 2,000-meter trench or on a shallow reef. IDW will incorrectly interpolate high biomass across these entirely different habitats.\
**The "Bullseye" Effect on Patchy Data:** Fisheries data, particularly for rockfish, is famously patchy and often "zero-inflated" (many hauls with zero catch, and a few hauls with massive catches). IDW handles extreme values poorly, smoothing them out in a circular radius that creates unrealistic "bullseyes" of high density around a single large haul, rather than reflecting actual habitat contours.\
**Assumes Isotropic Distribution:** IDW assumes that spatial influence spreads equally in all directions (isotropy). In marine shelf environments, habitats and oceanographic conditions typically stretch parallel to the coastline or shelf break. Therefore, fish distributions are highly directional (anisotropic).\
**No Environmental Covariates:** IDW is strictly mathematical and cannot incorporate driving environmental factors. It predicts fish presence based solely on where other fish were caught, rather than modeling why they are there based on variables like bottom temperature, substrate type, or depth. **Lack of Uncertainty Estimation:** IDW is a deterministic method, meaning it generates a fixed surface but provides no estimate of variance or error. For fisheries management and stock assessments, knowing the confidence intervals of a spatial distribution is just as important as the estimate itself.\
**Better Alternatives**\
For data like this, fisheries scientists typically use model-based approaches that account for environmental covariates and spatial autocorrelation. Vector Autoregressive Spatio-Temporal (VAST) models, Generalized Additive Models (GAMs), or Kriging with external drift are standard because they allow bathymetry and habitat types to constrain the interpolation, resulting in maps that actually reflect the species' ecology."\
so I guess it's not good to use that method for various reasons, so he used the plotting function and made points weighted by cpue. it's just really hard to see patterns that way, but is what it is

### Time series and other plots and tables for the base case run (MyOutput)

```{r}
setwd(run_base)
MyOutput = SS_output(dir = run_base,covar = Uncertainty,forecast=TRUE,printstats=FALSE)





```

### Fishery tables

```{r}
 source("C:/GitProjects/goa_rex/R/query_catch_blend.R", echo = TRUE)
catch_data<-query_catch_blend(akfin_conn=akfin,sp_group_code='REXS',region= c('WG','CG','WY','EY','SE'))

tot_catch_df<-catch_data %>% 
  dplyr::group_by(year) %>%
  dplyr::summarise(Catch = sum(weight_posted))
write.csv(tot_catch_df,file.path(DataDir,"Total_Catch.csv"))


r_d_catch_df<-catch_data %>% 
  dplyr::group_by(year,retained_or_discarded) %>%
  dplyr::summarise(catch = sum(weight_posted)) %>%
  tidyr::pivot_wider(id_cols=year,names_from = retained_or_discarded,values_from = catch) %>%
  dplyr::mutate(total_catch = D + R) %>%
  dplyr::rename(Discarded = D, Retained = R, Total = total_catch) %>%
  dplyr::mutate(Proportion_Retained = Retained/Total)
write.csv(r_d_catch_df, file.path(DataDir,"R_D_Catch.csv"))

catch_by_gear_df<-catch_data %>%
  dplyr::group_by(year,agency_gear_code) %>%
  dplyr::summarise(catch = sum(weight_posted)) %>%
  dplyr::left_join(tot_catch_df) %>%
  dplyr::mutate(prop = catch/Catch) %>%
  tidyr::pivot_wider(id_cols = year, names_from = agency_gear_code,values_from = prop) %>% dplyr::mutate(across(where(is.numeric), ~ replace_na(.x, 0)),
               non_pelagic_trawl = NPT + BTR) %>%
  dplyr::filter(year>=1996)
write.csv(catch_by_gear_df,file.path(DataDir,"catch_by_gear.csv"))

catch_by_season<-catch_data %>%
  dplyr::mutate(month = lubridate::month(week_end_date)) %>%
  dplyr::mutate(season = case_when(month <=3 ~ "Q1",
                                   month > 3 & month<=6 ~ "Q2",
                                   month >6 & month<9 ~ "Q3",
                                   TRUE ~ "Q4")) %>%
  dplyr::group_by(year,season) %>%
  dplyr::summarise(catch = sum(weight_posted)) %>%
  tidyr::pivot_wider(id_cols = year,names_from = season, values_from = catch) %>%
  left_join(tot_catch_df) %>%
  dplyr::mutate(Q1_prop=Q1/Catch, Q2_prop = Q2/Catch, Q3_prop = Q3/Catch, Q4_prop = Q4/Catch)
write.csv(catch_by_season, file.path(DataDir,"catch_by_season.csv"))


catch_by_nmfs_area_df<-catch_data %>%
  dplyr::group_by(year,reporting_area_code) %>%
  dplyr::summarise(catch = sum(weight_posted)) %>%
  dplyr::left_join(tot_catch_df) %>%
  dplyr::mutate(prop = catch/Catch) %>%
  tidyr::pivot_wider(id_cols = year,names_from = reporting_area_code, values_from = prop) %>%
  dplyr::mutate(across(where(is.numeric), ~ replace_na(.x,0)))

write.csv(catch_by_nmfs_area_df,file.path(DataDir,"catch_by_nmfs_area_df.csv"))

#just out of interest and could be combined with season, nmfs_area
catch_by_sector_df<-catch_data %>%
  dplyr::group_by(year,harvest_sector) %>%
  dplyr::summarise(catch = sum(weight_posted)) %>%
  tidyr::pivot_wider(id_cols = year,names_from = harvest_sector,values_from = catch)
write.csv(catch_by_sector_df,file.path(DataDir,"catch_by_sector.csv"))
```

### PSC tables

```{r}
#source the function
source("C:/GitProjects/goa_rex/R/query_psc.R", echo = TRUE)
psc<-query_psc(akfin_conn = akfin,region = 'GOA')

target_species <- psc %>% distinct(trip_target_name)
psc_all_species<-psc %>% 
  dplyr::group_by(year,species_group_name) %>%
  dplyr:: mutate(halibut_mortality_tons = tidyr::replace_na(halibut_mortality_tons, 0)) %>%
  dplyr::summarise(pscnq_estimate = sum(pscnq_estimate),halibut_mortality_tons = sum(halibut_mortality_tons))

psc_species<-psc %>%
  dplyr::filter(trip_target_name == "Rex Sole - GOA") %>%
    dplyr:: mutate(halibut_mortality_tons = tidyr::replace_na(halibut_mortality_tons, 0)) %>%
  group_by(year,species_group_name) %>%
  dplyr::summarise(pscnq_estimate_rex = sum(pscnq_estimate),halibut_mortality_rex = sum(halibut_mortality_tons)) %>%
  dplyr::left_join(psc_all_species) %>%
  dplyr::mutate(pscnq_prop = pscnq_estimate_rex/pscnq_estimate, halibut_mortality_prop = halibut_mortality_rex/halibut_mortality_tons,halibut_mortality_prop = tidyr::replace_na(halibut_mortality_prop,"")) %>%
  dplyr::select(year,species_group_name,pscnq_prop,halibut_mortality_prop)

pscnq_halibut<-psc_species %>%
  dplyr::filter(species_group_name == "Halibut")
write.csv(pscnq_halibut,file.path(DataDir,"psc_halibut_proportion_from_rex.csv"))

pscnq_other<-psc_species %>%
  dplyr::filter(species_group_name!="Halibut") %>%
  dplyr::select(-c(halibut_mortality_prop)) %>%
  tidyr::pivot_wider(id_cols = year,names_from = species_group_name, values_from = pscnq_prop) %>%
  dplyr::mutate(across(everything(), ~ round(.x, 3)),
                across(everything(), ~ replace_na(.x,0))) 
write.csv(pscnq_other,file.path(DataDir,"psc_other_proportion_from_rex.csv"))

```

### Survey tables

Here we need a table with survey biomass, stdev with depth included.

```{r}
surv_df<-get_biomass_by_depth(srv_sp_str=c('10200'),surv_id = c('47'),conn = akfin)

```

## Likelihood values

```{r}
base_like<-MyOutput$likelihoods_used %>% 
  dplyr::rename(model_25.1 = values) %>%
  dplyr::select(-lambdas) %>%
  tibble::rownames_to_column(var = "component")

all_like<-model_25.0$likelihoods_used %>% 
  dplyr::rename(model_25.0 = values) %>%
  dplyr::select(-lambdas) %>%
  tibble::rownames_to_column(var = "component") %>%
  dplyr::full_join(base_like)

write.csv(all_like,file.path(compare_dir,"compare_likelihoods.csv"))

```

## Parameter tables

Fs, rec devs, main parameters, selex parameters

param values for a single run (also is with Time_Series_Param_Tables_2_Growth_Morphs.R):

```{r}
library(stringr)

for (i in 1:length(the_run_list)) {
the_run_pars<-data.frame(Label = the_run_list[[i]]$parameters$Label, Value = the_run_list[[i]]$parameters$Value, Phase = the_run_list[[i]]$parameters$Phase, StDev = the_run_list[[i]]$parameters$Parm_StDev)

the_run_pars<-the_run_pars %>% mutate(model = the_run_labels[i])
if (i == 1) {
  base_pars_df <-the_run_pars
} else {
  base_pars_df<-base_pars_df %>% bind_rows(the_run_pars)
}
}
# new_data_pars_df<-data.frame(Label = the_run_list[i]$parameters$Label, Value = MyOutput$parameters$Value, Phase = MyOutput$parameters$Phase, StDev = MyOutput$parameters$Parm_StDev)
#   
base_pars_df<-na.omit(base_pars_df[base_pars_df$Phase >0,])

main_params_df<-base_pars_df %>% filter(!str_detect(Label, "^Early"),!str_detect(Label,"^Main"))

pars_df<-main_params_df %>% 
  mutate(area = case_when(str_detect(Label,"GP_2")==TRUE ~ "e",
                               str_detect(Label,"GP_1")==TRUE ~ "wc",
                               str_detect(Label,"Western") == TRUE ~"wc",
                               str_detect(Label,"Eastern") == TRUE ~ "e",
                               str_detect(Label,"Fishery") == TRUE ~ "wc",
                               TRUE ~ "global"),
         sex = case_when(str_detect(Label,"_Fem")==TRUE ~ "(f)",
                         str_detect(Label,"_Mal") == TRUE ~"(m)",
                         TRUE ~ "global")) %>%
 # filter(!str_detect(Label,"Age")) %>% 
  select(-c(Phase)) %>% 
  mutate(Label = gsub("_GP_1", "", Label),Label = gsub("_GP_2","",Label),Label = gsub("_Fem","",Label),Label = gsub("_Mal","",Label),
         Label= case_match(Label,"NatM_uniform" ~ "Natural mortality",
                                 "L_at_Amin" ~ "Length at age 2",
                                 "L_at_Amax" ~ "Length at age 20",
                                 "VonBert_K" ~ "Von Bertalanffy k",
                                 "CV_young" ~ "CV in length at age 2",
                                 "CV_old" ~ "CV in length at age 20",
                                 "RecrDist_area_2_month_1" ~ "Recruitment proportion to Eastern GOA",
                           "SR_LN(R0)" ~ "Mean recruitment (log)",
                           "LnQ_base_WesternCentralSurvey(2)" ~ "Catchability (log)",
                           .default = Label)) 
                           
names(pars_df) <-c("parameter","est","stdev","model","area","sex")

area_pars<-pars_df %>% filter(area !="global", sex!="global") %>%
  mutate(parameter = paste(parameter,sex)) %>%
  select(-sex) %>%
  pivot_wider(id_cols = parameter,names_from = c(model,area),values_from = c(est,stdev)) 

write.csv(area_pars,file.path(compare_dir,"area_specific_pars.csv"))


#wc_bio_pars_df<-pars_df %>% filter(area == "wc")
#e_bio_pars_df <-pars_df %>% filter(area == "e")
global_bio_pars_df<-pars_df %>% filter(area == "global" | sex == "global") %>%
  relocate(parameter) %>% 
  select(-c(area,sex)) %>%
  tidyr::pivot_wider(id_cols = parameter, names_from = model, values_from = c(est,stdev))
write.csv(global_bio_pars_df,file.path(compare_dir,"global_and_selex_pars.csv"))


RecPars.df<-data.frame(type = MyOutput$recruitpars$type, Yr = MyOutput$recruitpars$Yr, Value = MyOutput$recruitpars$Value, StDev = MyOutput$recruitpars$Parm_StDev)

write.csv(RecPars.df,file = paste0(SS3PredDir,"RecPars.csv")) 
write.csv(EstP.df,file = paste0(SS3PredDir,"Main_Pars.csv")) #easiest with fishing mortality to copy and paste straight from the report file.
```

### Time Series Tables

Mostly I ran this file: "C:/GitProjects/newsbss/Process_SS_Output/Time_Series_Param_Tables_2_Growth_Morphs.R"

```{r biomass_by_area_table}
c<-MyOutput


#area-specific total and spawning biomass
gp2_bio_df<-c$timeseries %>% 
  dplyr::filter(Yr>=FirstYr, Yr<=LastYr, Area==2) %>%
  dplyr::select(Yr,'SmryBio_SX:1_GP:2','SpawnBio_GP:2')

gp_bio_df<-c$timeseries %>% 
  dplyr::filter(Yr>=FirstYr, Yr<=LastYr, Area==1) %>%
  dplyr::select(Yr,'SmryBio_SX:1_GP:1','SpawnBio_GP:1') %>%
  full_join(gp2_bio_df) %>%
  rename(Year = Yr,tot_bio_wc = 'SmryBio_SX:1_GP:1',ssb_wc = 'SpawnBio_GP:1',tot_bio_e = 'SmryBio_SX:1_GP:2', ssb_e = 'SpawnBio_GP:2')


```

### Troubleshooting projections

The models are nearly identical and the projections are quite different, so something is going wrong. Here are the experiments I'm doing:\
No: 1. Ran the new base case (run_19) with an old version of ss3 and then doing projections with current spmR to see if something changed in Report.sso that has impacted the data input file for spmR

No: 2. I had changed summary biomass to start at age 3 for this years runs. change it back. If this works take a close look at the ForProjections_GrowthMorth1.dat file to make sure it's right with recruitment occurring at age 3.

*3. New ss3 and new spm with 2021 model (this was a problem - it yields a huge B0 as compared to 2021 and 2024 projections)*

Next, I looked closely at the .dat file to see if it was right

4.  Run old spm from 2021 with new base model (not easy because spm.exe was in the environment path on old computer and going further back I was running main.exe, and so couldn't run my make_exec function with that - too many changes in output organization)

5.  *Cut recruitment in the projection .dat file in half, based on memory that at some point in the change to main/proj to spm versions, the spm.tpl code changed to assume that only female recruitment would be input, and it also assumes equal sex ratios at age of recruitment (here it's age 3), so it multiplies input recruitment by 2.*

6.  Check current spm.tpl to make sure I'm right about (5). Yes, in spm.tpl, lines 630-634 in the PARAMETER_SECTION under LOCAL_CALCS (and you can see that the earlier assumption is commented out, and that this applies to 2-sex model output.

    ```         
    if (nsexes(ispp)==1) 
      AMeanRec(ispp) = mean(R(ispp));  // Arithmetic mean
    else
      AMeanRec(ispp) = 2.*mean(R(ispp));  // Arithmetic mean, converted to females (half) later
      // AMeanRec(ispp) = mean(R(ispp));  // Arithmetic mean, converted to females (half) later
    ```

The solution is to fix this in the AreaGrowthRex package under R/write_proj_ss_2Area.R and R/write_proj_ss_1Area.R by either only entering female recruits or by dividing in half prior to writing it out.

The following troubleshooting runs for projections required running the little chunks below and then going further up in the code to @organize_projections to set the_proj_dirs to each of these dirs, and running that @organize_projections and then subequently @make_exec_summary_tables, pointing to the usefulness of making those chunks into functions for next time.

```{r troubleshoot projections}
base_old_ss3_dir<-file.path(run_dir,"run_19_old_ss3")

base_old_ss3_output = SS_output(dir = base_old_ss3_dir,covar = TRUE,forecast=TRUE,printstats=FALSE)

AreaGrowthRex::get_projections(mydir = base_old_ss3_dir ,MyOutput = base_old_ss3_output,NagesProject = 20,FirstYr = 1982,LastYr = 2021,RecruitmentAge=3,species="Rex")
SS_plots(replist = base_old_ss3_output,uncertainty=TRUE) 

```

```{r troubleshoot summary biomass}
age0_smry_bio_dir<-file.path(run_dir,"run_19_smry_biomass_age_0")
age0_smry_bio_output<-SS_output(dir = age0_smry_bio_dir,covar = TRUE,forecast=TRUE,printstats=FALSE)

AreaGrowthRex::get_projections(mydir = age0_smry_bio_dir ,MyOutput = age0_smry_bio_output,NagesProject = 20,FirstYr = 1982,LastYr = end_year,RecruitmentAge=3,species="Rex")
```

```{r troubleshoot 2021 projections}
run_2021_new_ss3_dir<-file.path(run_dir,"run_2021_accepted_new_ss3v")
run_2021_new_ss3_output<-SS_output(dir = run_2021_new_ss3,covar = TRUE,forecast=TRUE,printstats=FALSE)

AreaGrowthRex::get_projections(mydir = run_2021_new_ss3_dir ,MyOutput = run_2021_new_ss3_output,NagesProject = 20,FirstYr = 1982,LastYr = end_year,RecruitmentAge=3,species="Rex")
```

```{r troubleshoot spm version}
run_19_old_spm_dir<-file.path(run_dir,"run_19_old_spm")
run_19_old_spm_output<-SS_output(dir = run_19_old_spm_dir,covar = TRUE,forecast=TRUE,printstats=FALSE)

AreaGrowthRex::get_projections(mydir = run_19_old_spm_dir,MyOutput = run_19_old_spm_output,NagesProject = 20,FirstYr = 1982,LastYr = end_year,RecruitmentAge=3,species="Rex")
```

```{r divide recruitment in half}
run_19_half_rec_dir<-file.path(run_dir,"run_19_divide_rec_in_half")
run_19_half_rec_output<-SS_output(dir = run_19_half_rec_dir,covar = TRUE,forecast=TRUE,printstats=FALSE)

```
